### 2021-01-10 paper/main.tex
Initial commit of in-progress NSR20 paper files

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

% Additional Packages:
\usepackage{graphicx}
\usepackage[labelformat=simple]{subcaption}
\renewcommand\thesubfigure{(\alph{subfigure})}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{nicefrac}
\usepackage{textcomp} % For \textdegree (also try \usepackage{gensymb} which gives \degree)
\usepackage{soul} % For highlighting text

% Macro Definitions:
\def\s{\sigma}
\providecommand{\e}[1]{\ensuremath{\times 10^{#1}}}

\title{Automated data-driven symbolic modeling of the lift forces acting on flapping wings}

\author[1,*]{Charles Richter}
\author[2,3]{Hod Lipson}
% \author[1,2,+]{Christine Author}
% \author[2,+]{Derek Author}
\affil[1]{Cornell University, Sibley School of Mechanical and Aerospace
Engineering, Ithaca, 14850, United States}
\affil[2]{Columbia University, Department of Mechanical Engineering, New York,
10027, United States}
\affil[3]{Columbia University, Data Science Institute, New York, 10027, United
States}

\affil[*]{car45@cornell.edu}

% \affil[+]{these authors contributed equally to this work}

\keywords{flapping flight, unsteady aerodynamics, data mining, symbolic
regression, evolutionary computation}

\begin{abstract}
We explore a data-driven approach to symbolically modeling the lift produced by
the flapping wing of a micro air vehicle.  The aerodynamics of flapping flight
pose a modeling challenge due to complex effects related to vortex dynamics,
wake interaction, aeroelasticity and three-dimensional flow, yet simple and
accurate equations are useful for engineering purposes and optimization of wing
design.  Experiments were conducted to measure the flapping kinematics and
instantaneous lift forces produced by eleven different 3D printed wing
geometries at three different flapping amplitudes and multiple speeds, totaling
68 distinct sets of data.  58 of these data sets were used to train data-driven
models using a symbolic regression algorithm, while the remaining sets were
reserved for model testing.  Data-driven models were found to be more accurate
and much simpler than existing quasi-steady analytical models consistent with
aerodynamic theory.  Furthermore, data-driven models were found to separate
distinct physical effects into inertial and aerodynamic components bearing
remarkable resemblance to their analytical counterparts. This result suggests
that a data-driven approach could be a powerful tool in separating and
identifying the structural building blocks of a wide variety of unknown
dynamical systems.
\end{abstract}
\begin{document}

\flushbottom
\maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
\thispagestyle{empty}

% \noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.

\section*{Introduction}

% The Introduction section, of referenced text\cite{Figueredo:2009dg} expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.

Forces produced by the flapping wings of insects, birds and micro air vehicles
are of crucial importance for flight, yet they are subtle and difficult to
model.  Many studies have developed equations and solution techniques to
understand the fluid flow around these wings, to compute control actions
required for stable flight, or to optimize wing design. Aerodynamic studies
related to insect flight and intermediate Reynolds numbers (Re 100-10,000) have
generally focused on either numerical solutions of Navier-Stokes equations in
two or three dimensions \cite{wu2004unsteady,wang2000two}, quasi-steady
approximations of the observed behavior
\cite{sane2002aerodynamic,ellington1984aerodynamics}, or a comparison of both
methods \cite{pesavento2004falling}.  The Navier-Stokes equations offer
important insights into the full character of the fluid around the wing
including vorticity and fluid-structure interaction, but are complex and
computationally expensive to calculate.  Quasi-steady approximations, in
contrast, capture most of the character of the aerodynamic forces without any
detailed modeling of the fluid flow.  These equations are phenomenological in
nature and constitute an assembly of terms that are consistent with aerodynamic
theory that can be used to represent lift, drag and other forces, as functions
of a wing's kinematic variables such as velocity and angle of attack.  Both
methods have shown qualitative and quantitative agreement with experimental
measurements \cite{wang2004unsteady}.

% \subsection{Automated modeling}

Quasi-steady representations of lift are particularly well suited to the process of symbolic
regression, which harnesses computational power to conduct an automated search over the space of all
kinematic variables and geometric parameters to find equations that most accurately predict the
forces on a wing.  Symbolic regression algorithms are based on evolutionary computation
\cite{koza1992genetic,forrest1993genetic} and have been used to model complex systems defined by explicit
\cite{duffy2002using,fillon2007symbolic,bautu2005symbolic}, implicit \cite{schmidt2010symbolic}, iterated
\cite{schmidt2009solving} and differential equations \cite{bongard2007automated}.  The
process underlying this evolutionary computation approach is to generate an initially random
population of equations combining time-series data and constant variables with algebraic operators,
analytic functions and constants.  These candidate equations are then evaluated to measure their
accuracy in predicting the measured output of the system or partial derivatives within the data.  In
each iteration of the algorithm, the best candidates are retained and varied by recombining and
varying their subexpressions. The algorithm is terminated when a desired level of accuracy is
reached.  In this case, the time-series data includes the orientations, velocities and accelerations
of wings, the constant variables include chord lengths, wingspans, centroids of wing area and other
physical constants, and the known output of the system is the measured vertical force produced
during flapping.  The resulting equations returned by the algorithm are quasi-steady models that can
be compared to existing equations from the literature.

% \subsection{Measurement of lift forces}

Numerous studies have measured the forces produced by a flapping wing in
various experimental setups including flapping in a liquid bath such as oil
\cite{walker2009photogrammetric}, by tracking the kinematics of insects in
flight \cite{fry2003aerodynamics,ristroph2009automated} and by tracking the
detailed wing motions and visualizing the airflow around the wings of insects
in a wind tunnel \cite{young2009details}.  Only recently have studies directly
measured the forces involved in flapping with at-scale robotic wings or actual
insects in air \cite{whitney2010aeromechanics,graetzel2008real}.  In general,
measuring the forces produced by flapping wings in air is difficult because the
magnitude of the forces is small, requiring custom-built capacitive or MEMS
force sensors, and because the speed of flapping and prominence of inertial
effects cause significant vibrations that obscure the force measurement.  The
flapping wings examined in this work represent a size range appropriate for
micro air vehicles, and produced forces large enough to be measured by a
commercially available load cell.

Here we present a series of experiments measuring the kinematics and lift forces produced by eleven
differently sized wings flapping at multiple speeds and amplitudes with Reynolds numbers between
2500 and 5000. Several quasi-steady models from the literature are shown to exhibit strong agreement
with the experimental data.  Then a set of data-driven equations were automatically generated from
the data using the symbolic regression algorithm developed by Schmidt and Lipson \cite{schmidt2009distilling}.
These data-driven models exhibit several remarkable properties.  First, they are at least as
accurate as quasi-steady equations from the literature yet they are much simpler.  Second, they
naturally separate distinct, understandable physical effects including terms corresponding to
inertial forces and aerodynamic lift.  Finally, it is shown that models generated from data of a
single experiment do not have the property of separating distinct physical effects, and they fail to
accurately predict forces produced by a wing of a different geometry or flapping pattern.  It is
only when models are trained on a large set of experiments including many different wings and
flapping patterns that over-fit candidate equations are discarded and those that reflect the true
underlying physics are maintained.  The models trained on a diverse set of data are shown to
accurately predict the forces produced by wings of similar size and flapping kinematics that were
not included in the training set. The ability of the symbolic regression process to infer distinct
physical contributions to to the total lift force based on data alone suggests it may be a powerful
tool in identifying the underlying structure of a wide variety of complex systems.

\section*{Results}

% Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.
%
% \subsection*{Subsection}
%
% Example text under a subsection. Bulleted lists may be used where appropriate, e.g.
%
% \begin{itemize}
% \item First item
% \item Second item
% \end{itemize}
%
% \subsubsection*{Third-level section}
%  
% Topical subheadings are allowed.

\subsection{Data-driven models}
The strategy for modeling lift forces symbolically was to train candidate equations on 58 of the 68 total experiments simultaneously, with the remaining ten reserved for testing.  These 58 training sets were intended to represent a diverse sampling of wing geometry, flapping speed and amplitude combinations.  The results from these 58 experiments were compiled into a single dataset listing roughly two periods of smoothed, registered flapping data per experiment.  The data consisted of instantaneous measured vertical force along with the measured value of each kinematic variable, the first and second derivatives of those variables, and the geometric parameters necessary to describe the wing's shape and distribution of area, totaling 31 separate variables.

By training on 58 experiments simultaneously, the symbolic regression algorithm generated equations that fit all of the data, using the some, but not all, of the provided variables to account for variations in kinematics and wing geometry between experiments.   Table~\ref{table:eqmodels} lists five of the leading candidate models, which were trained using mean absolute error (MAE) as the fitness metric during the search. The search was conducted for approximately 24 hours, using 60 cores on a computing cluster.

\begin{table}[ht]
\centering
\begin{tabular}{|l|l|r|r|r|}
\hline
  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & {\bf MAE (N)}\\
\hline
\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 1.19\e{-3}\\
\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 1.18\e{-3}\\
\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 1.14\e{-3}\\
\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 1.14\e{-3}\\
\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 1.12\e{-3}\\
\hline
\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.46\e{-3}\\
\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.28\e{-3}\\
\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.28\e{-3}\\
Whitney-Wood & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.21\e{-3}\\
Pesavento-Wang & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.20\e{-3}\\
\hline
\end{tabular}
\caption{\label{table:eqmodels}Comparison of data-driven models and analytical equations of lift.
The models are sorted by equation size, which is calculated as the sum of arithmetic and
trigonometric operators required to compute a force prediction from kinematic and geometric data.
MAE reports average mean absolute error (in Newtons of vertical force) for the ten test
experiments.}
\end{table}

\subsection{Analytical models for comparison}
The analytical models used for comparison are also listed in Table~\ref{table:eqmodels}.  These models consist of calculated inertial forces added to a collection of aerodynamic terms.  The primary aerodynamic contribution in all quasi-steady models is a translational lift term proportional to the squared velocity of each span-element of the wing that is related to the wing's angle of attack and empirically determined coefficient of lift.  This term is labeled $L_{planar}$ when wing motions out of the nominal horizontal flapping plane are neglected in the calculations, or $L_{trans}$ when out-of-plane motions are included. This translational lift term is given by:
\begin{equation}
L_{trans}=(\nicefrac{1}{2})\rho v^2AC_L,
\end{equation}
where \(v\) is the wing velocity, \(\rho\) is the fluid density, \(A\) is the area of the wing surface, and \(C_L\) is the coefficient of lift.  The theoretical value of the coefficient of lift depends on the angle of attack \(\alpha\):
\begin{equation}
C_L = C_{Lmax}\sin(2\alpha).
\end{equation}

Other aerodynamic contributors include rotational lift proportional to the product of a wing's rotational and translational velocities, labeled $L_{rot}$, as well as "added-mass" effects related to the accelerations of fluid surrounding the wing \cite{sedov1965two}.  These added mass terms are proportional to the wing's linear acceleration ($AM_{trans}$), rotational acceleration ($AM_{rot}$), or a product of the two ($AM_{trans-rot}$).  Two sets of quasi-steady equations from the literature include various combinations of such terms \cite{whitney2010aeromechanics,pesavento2004falling}, which allow more flexible model fitting through additional coefficients.

Inertial forces acting on the wings are modeled as a sum of two contributions.  The primary contribution results from the pendulum-like swinging motion that occurs when the deflection angle ($\psi$) flips at stroke reversal.  The secondary contribution results from small angle deviations from the horizontal flapping plane ($\theta$).  These angles are depicted in Figure~\ref{fig:apparatus}.  These inertial forces are given by:
\begin{equation}\label{sdbla}
F_{pendulum} = Y_{CM}M(\ddot{\psi}\sin\psi + \dot{\psi}^2\cos\psi)
\end{equation}
and
\begin{equation}\label{sdbla2}
F_{deviation} = -X_{CM}M(\ddot{\theta}\cos\theta - \dot{\theta}^2\sin\theta),
\end{equation}
\noindent where \(X_{CM}\) and \(Y_{CM}\) are the coordinates of the center of wing area and \(M\) is the mass of the wing surface.

\begin{figure}%[!htb]
\centering
\begin{subfigure}{0.3\textwidth}
\includegraphics[trim=0 15 0 5, clip, width=\textwidth]{figures/apparatus}
\caption{\label{fig:apparatus}}
\end{subfigure}
\qquad
\begin{subfigure}{0.5\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/pareto_mae_nofit_PNAS}
\caption{\label{fig:pareto_mae}}
\end{subfigure}
\caption{\subref{fig:apparatus} Schematic diagram of the experimental apparatus showing three
  kinematic variables of the wing motion: stroke ($\phi$), deflection ($\psi$) and deviation
  ($\theta$). \subref{fig:pareto_mae} Pareto plot of all data-driven (``EQ Models'') and analytical
  models using mean values for model coefficients. All data-driven models were smaller and more
  accurate than analytical models.
}
\end{figure}

\subsection{Evaluation of Model Performance}
An evaluation of model performance allows comparison between analytical models and the equations developed through the data mining process.  Equations are judged by their accuracy and complexity.  In this case, the complexity or size of an equation was measured by simply counting the number of arithmetic and trigonometric operators used to combine the measured kinematic variables and geometric properties into an equation for instantaneous vertical force.  Both our data-driven models and the analytical equations are measured in this way.  For fairness, constants in the analytical models that would have been lumped together into larger coefficients by the symbolic regression algorithm (such as material density, air density, \(\pi\) or factors of integration) are also lumped together in the calculations of size for the analytical equations.  The size of the inertial calculations was added to the sizes of analytical lift calculations.  The 
pendulum motion has an equation size of 13 and the total inertial calculation including out-of-plane deviation has a size of 27.

Table~\ref{table:eqmodels} shows the size of all equations considered in this study, and indicates that the analytical models are all significantly more complex than our data-driven models.  One of the reasons for high complexity in the analytical models is the dependence on wing geometry and integration of forces over the span and chord of the wing.  A second reason for the complexity in analytical models is the number of geometric and coordinate transformations required to compute the necessary linear and rotational velocities.  In this sense, several terms generated during symbolic regression can be considered to be approximations of much more subtle calculations of kinematic quantities.

The accuracy of each equation was measured as the mean absolute error (MAE) between the predicted and actual vertical force (in Newtons) produced by the flapping wing.  Figure~\ref{fig:pareto_mae} shows the Pareto plot of all models considered, depicting accuracy as a function of model complexity. It is important to note that all data-driven models have between three and five coefficients, so tuning these coefficients allows further minimization of error compared to the analytical models, which have between one and three.  Nevertheless, discovery of models with smaller error than theoretical predictions that are simultaneously much simpler is a significant feat for the data mining process and symbolic regression.

\begin{figure}[ht]
\centering
\includegraphics[width=\textwidth]{figures/eq_goodbad_PNAS}
\caption{Comparison of force predictions between $EQ_4$ (blue) and Equation~\eqref{eq:badeq} (red).
  $EQ_4$ was trained on 58 distinct experiments simultaneously, including the data shown in (A),
  whereas Equation~\eqref{eq:badeq} was trained \emph{only} on the data shown in (A).  Panel (A)
  shows that both models perform well on this data within their training sets.  However, when
  presented with data outside their training set (B,C,D), only $EQ_4$ is able to generalize and
  provide accurate predictions, while Equation~\eqref{eq:badeq} fails to do so.}
\label{fig:eq_goodbad}
\end{figure}

\subsection{Performance of models trained on one experiment versus many experiments}
Figure~\ref{fig:eq_goodbad}(A) illustrates the predictive performance of two models on data within their training sets, while Figure~\ref{fig:eq_goodbad}(B,C,D) illustrate the performance of those same two models on data outside their training sets. The differences between these two models result from the fact that one of them, $EQ_4$ from Table~\ref{table:eqmodels} (drawn in blue), was trained on 58 separate experiments simultaneously, whereas the other one, shown below (drawn in red), is one of many models trained only on the single experiment shown in Figure~\ref{fig:eq_goodbad}(A):
\begin{align}
 Force = \frac{C_1 + C_2\psi^2 - C_3\psi}{C_4 + \ddot{\theta} + C_5\dot{\theta} + \cos(C_6 -
  \frac{C_7}{\psi^2} - C_8\psi)} - C_9.
\label{eq:badeq}
\end{align}
Thus, $EQ_4$ shows strong qualitative and quantitative agreement with all of the experiments in Figure~\ref{fig:eq_goodbad}, indicating that this model is not over-fit but in fact generalizes well to new data. Equation~\eqref{eq:badeq}, on the other hand, fails to accurately predict the recorded data outside its training set.  Insufficient data and over-fitting are potential pitfalls of data-driven modeling, but are overcome in this case with a diverse sampling of training data.

Since Equation~\eqref{eq:badeq} is unable to accurately predict forces outside its training set, it is unlikely to offer any insight into the underlying physical processes.  $EQ_4$, however, must capture some of the underlying structure of the physics in order to generalize to data upon which it was not trained.  In data-driven modeling, training on many experiments simultaneously helps to reject candidate equations that may fit some, but not all, of the data, and helps to retain those models that reflect the underlying physics and offer insight into the structure of the additive components that compose the observed behavior of a complex dynamical system.

\subsection{Correspondence of physical terms}
One powerful result of the data-driven modeling approach is that in many cases, certain terms from the data-driven models closely match physically understandable terms from the analytical models.  The correspondence between these terms is listed in Table~\ref{table:eqcorrespondence}. Taking model $EQ_4$ as an example, all three terms show strong resemblance to inertial and aerodynamic components of the analytical models.  Figure~\ref{fig:eqcomparison} shows a comparison between the simple velocity-squared aerodynamic lift prediction and the second term of $EQ_4$: \(C_2RX_{CM}\psi\dot{\phi}\).  These two calculations are numerically very similar, yet the third term of $EQ_4$ is much simpler than $L_{trans}$.

Similarly, the sum of terms 1 and 3, \(C_1RY_{CM}\dot{\psi}^2\cos(\psi) - C_3R\ddot{\theta}\), closely approximates the calculation of inertial forces including the pendulum motion of the wing as well as the out-of-plane deviation.  Figure~\ref{fig:eqcomparison} also shows these two calculations, indicating very similar structure.  While $EQ_4$'s approximation of these inertial forces does not precisely match the analytical calculation, it follows a very similar pattern and is much simpler.  Remarkably, these inertial terms utilize some of the same variable combinations as the analytical calculations to express the inertial force, namely \(Y_{CM}\dot{\psi}^2\cos(\psi)\) and \(\ddot{\theta}\), with different coefficients to express the geometric variation between wings and small angle approximations \(\cos\theta\approx1\), \(\sin\theta\approx0\) for small out-of-plane deviation angles.  Since \(|\theta|<2^{\circ}\) in general, the small angle approximation is valid.

\begin{table}[ht]
\centering
\begin{tabular}{|l|l|l|}
\hline
{\bf Contribution} & {\bf Analytical} & {\bf Data-Driven ($EQ_4$)}\\
\hline
Aerodynamic Lift & \(\nicefrac{1}{2}\rho v^2AC_{Lmax}\sin(2\alpha)\) & \(C_2RX_{CM}\psi\dot{\phi}\)\\
Pendulum 1 & \(M\cdot Y_{CM}\ddot{\psi}\sin{\psi}\) & \it{None} \\
Pendulum 2 & \(M\cdot Y_{CM}\dot{\psi}^2\cos{\psi}\) & \(C_1R\cdot Y_{CM}\dot{\psi}^2\cos{\psi}\) \\
Deviation 1 & \(\approx-MX_{CM}\ddot{\theta}\) & \(- C_3R\ddot{\theta}\)\\
Deviation 2 & \(\approx0\) & \it{None} \\
%Deviation 2 & \(X_{CM}M\dot{\theta}^2\sin\theta\) & \(<none>\)\\
\hline
\end{tabular}
\caption{\label{table:eqcorrespondence}Correspondence between analytical and data-driven terms.}
\end{table}

\begin{figure}[ht]
\centering
\includegraphics[width=.48\textwidth]{figures/eq_comparison_PNAS}
\caption{Comparison between terms of the analytical expressions (red) and data-driven models (blue).
  (A) Primary translational aerodynamic lift proportional to the square of the wing velocity. (B)
  Inertial forces related to the pendulum motion of the wing and deviation out of the stroke plane.
  The modeling process automatically decoupled aerodynamic and inertial effects with no prior
  knowledge of the system.  The total measured force is shown for reference (black).}
\label{fig:eqcomparison}
\end{figure}

\section*{Discussion}

% The Discussion should be succinct and must not contain subheadings.

Table~\ref{table:eqcorrespondence} lists the corresponding terms between the
basic analytical expressions and the data-driven model $EQ_4$.  While two of
the inertial terms share the same combinations of variables, the data-driven
expression for aerodynamic lift is very different from its analytical
counterpart, and one portion of the pendulum motion is omitted altogether from
the data-driven models.  The reasons for these differences lie in the nature of
the symbolic regression algorithm, which is designed to return only the most
parsimonious expressions capable of predicting the experimental data at a given
level of accuracy.  Therefore, key physical effects may be represented
numerically by terms that are symbolically different and often much simpler
than their analytical counterparts.  This behavior of automatically separating
distinct dynamical building blocks is one of the most powerful aspects of
symbolic regression.  By sweeping the range of several kinematic and geometric
parameters in the 58 sets of training data, any candidate models that
incorrectly mix or couple two separate physical effects are discarded upon
testing on a data set with different parameters.

Nevertheless, in some domains such as flapping flight, there exist high degrees
of correlation between certain variables.  For example, the center of area
location in a rectangular wing is closely related to its span and chord length.
Similarly, kinematic variables such as stroke and deflection angles are highly
correlated with each other and their derivatives across all 58 experiments,
making these variables approximately interchangeable with each other when
coefficients are adjusted accordingly.  This behavior provides an explanation
for why \(\psi\dot{\phi}\) is used as an approximation for an aerodynamic lift
term proportional to \(v^2\).  These correlations between variables are
inherent in highly symmetric, periodic phenomena such as flapping flight.  In
principle, additional experiments could be performed on flapping surfaces that
are not bound to the natural kinematics of flapping wings to explore the full
parameter space and disambiguate non-physical terms in the candidate models.

% \section{Conclusions} 
The major thrust of this work was to apply a data-driven approach to modeling
the forces on a flapping wing.  The models produced by our symbolic regression
algorithm were shown to be more accurate than any of the analytical models, and
were much simpler, in some cases by an order of magnitude in equation size.
Furthermore, by dissecting these equations and examining individual terms, it
was observed that the data-driven models separated the measured forces into
understandable physical effects.  A lift term was found that was numerically
similar to the standard translational lift, and an inertial term was found that
was numerically similar to the complete calculation of the pendulum and
out-of-plane motions with some analytical terms in common.  Finally, by
training data-driven models on experimental data from 58 individual
wing-amplitude-speed combinations simultaneously, the symbolic-regression
algorithm was able to discard models that did not generalize, while retaining
those whose terms captured the underlying physical processes.

This application of a data-driven approach suggests that symbolic regression
could be useful in phenomenological modeling and guiding the exploration of a
wide variety of complex systems.  Its ability to automatically identify
distinct physical effects that are consistent across all datasets makes it a
powerful tool for identifying the structure of contributions from building
blocks that make up the complete observed behavior of a dynamical system.

\section*{Methods}

% Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.

\subsection{3D Printed wings and actuation mechanism}
The experimental wings were fabricated using an Objet Connex500 3D printer with
the rigid Objet FullCure720 material.
Flapping flight with 3D printed wings was reported in \cite{richter2011untethered}. The
wing design consisted of an interlocking root and tip, allowing the wing
surface to pivot naturally during flapping from its neutral vertical position
to an angle of attack of approximately 45\textdegree. The wing design is illustrated in
Figure~\ref{fig:wing1} and the hinge design is
illustrated in Figure~\ref{fig:root2}.  In total, eleven wings were used,
ranging in wingspan from 40 mm to 120 mm and ranging in chord length from 40 mm
to 80 mm. Flapping amplitudes were constrained by the geometry of the wing root
design to be set at 72\textdegree, 81\textdegree, and 96\textdegree, and each
wing was flapped at speeds ranging between 1 Hz and 3.5 Hz for each of these
amplitude settings.  68 sets of data were recorded, representing Reynolds
numbers between 2500 and 5000.

% See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
\begin{figure}[h]%[!htb]
\centering
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=0.95\textwidth]{figures/wing1}
    \caption{\label{fig:wing1}}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[trim=0 105 0 0, clip, width=0.95\textwidth]{figures/root2}
    \caption{\label{fig:root2}}
\end{subfigure}
\caption{Detail of 3D printed wing \subref{fig:wing1} and wing root \subref{fig:root2} design
  showing concentric-cylindrical design to allow passive pivoting of the wing surface.}
\end{figure}

\subsection{Measurement apparatus}
The apparatus used to measure the forces and kinematics of flapping wings was
based on a FUTEK LSM250 parallelogram load cell with a USB interface for data
logging.  This force measurement system was capable of a sampling rate of 1 kHz
and a maximum measurement precision of 1\e{-9} N.  Mounted directly to the load
cell was a wing-flapping mechanism, which consisted of a DC gear motor mated to
a crank and connecting rod used to drive the experimental wings in a roughly
sinusoidal flapping motion. This system is pictured in
Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency
of the load cell and wing driving mechanism was roughly 70 Hz, the natural
frequency for the first bending mode of the experimental wing was approximately
26 Hz, and experimental wings were flapped at or below approximately 3.5 Hz.
All experimental force measurements were filtered using a high-order zero-phase
digital low-pass filter with a cutoff frequency of 7.5 times the flapping
frequency, which preserved the primary inertial and aerodynamic forces acting
on the wings while eliminating the faster vibrations of the wing structure and
the load cell apparatus.

% % See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
\begin{figure}%[!htb]
\centering
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[height=5cm]{figures/apparatus2}
    \caption{\label{fig:apparatus2}}
\end{subfigure}
\begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[height=5cm]{figures/apparatus3}
    \caption{\label{fig:apparatus3}}
\end{subfigure}
\caption{ \subref{fig:apparatus2} Front view of the flapping wing apparatus showing the wing mounted
  on its driving mechanism attached to the force sensor. \subref{fig:apparatus3} Close-up view of
  the 3D printed frame and crank mechanism used to drive the flapping motion of the wing.}
\end{figure}

\subsection{High-speed measurement of flapping kinematics}
A Casio EXILIM EX-FH20 color video camera was used to record each experiment at
210 frames per second from a top-down perspective as well as a side view
provided by a mirror mounted in the field of view.  The main beam of each wing
was colored red to enable precise tracking of the stroke angle.  The top-down
projected area of the wing was used to extract deflection angle, and tracking
of the wing tip in the side-view was used to extract the angle of deviation out
of the wing's stroke plane.  These three angles precisely define the
configuration of the wing.  A schematic diagram of this apparatus is shown in
Figure~\ref{fig:apparatus}. A variety of image-processing techniques were used
to extract the kinematics of the wing motion from high-speed video.  The
red-colored main beam was identified with a color filter and a line was fitted
using a Hough Transform to identify the stroke angle.  The wing deflection
angle was computed by measuring the projected area of the wing in the image and
compared with the known dimensions of the rectangular wing shape. The
out-of-plane deviation angle of the wing was measured by tracking the leading
and trailing edge corners of the wingtip in the side-view of the mirror. We refer the reader to our
previous work for a more detailed description of our methods of extracting wing kinematics from
high speed video and processing the measured data\cite{richter2012thesis}.
Figure~\ref{fig:kinematics} provides sample force data and measured wing kinematics from one
experiment.

% \begin{figure}[ht]
% \centering
% \includegraphics[width=\linewidth]{figures/stream}
% \caption{Legend (350 words max). Example legend text.}
% \label{fig:stream}
% \end{figure}

\begin{figure}
\centering
\begin{subfigure}[b]{.45\linewidth}
\includegraphics[width=\linewidth]{figures/regforce1}
\caption{Measured vertical force.}
\label{fig:regforce}
\end{subfigure}
\begin{subfigure}[b]{.45\linewidth}
\includegraphics[width=\linewidth]{figures/regstroke1}
\caption{Wing stroke angle.}
\label{fig:regstroke}
\end{subfigure}

\begin{subfigure}[b]{.45\linewidth}
\includegraphics[width=\linewidth]{figures/regdef1}
\caption{Wing deflection angle.}
\label{fig:regdef}
\end{subfigure}
\begin{subfigure}[b]{.45\linewidth}
\includegraphics[width=\linewidth]{figures/regdev1}
\caption{Wing deviation angle.}
\label{fig:regdev}
\end{subfigure}
\caption{Measured \subref{fig:regforce} vertical force, \subref{fig:regstroke} wing stroke angle,
  \subref{fig:regdef} deflection angle, \subref{fig:regdev} and deviation angle from one experiment,
  registered and averaged over 18 flapping periods.}
\label{fig:kinematics}
\end{figure}

\bibliography{main}

% \noindent LaTeX formats citations and references automatically using the bibliography records in your .bib file, which you can edit via the project menu. Use the cite command for an inline citation, e.g.  \cite{Hao:gidmaps:2014}.
%
% For data citations of datasets uploaded to e.g. \emph{figshare}, please use the \verb|howpublished| option in the bib entry to specify the platform and the link, as in the \verb|Hao:gidmaps:2014| example in the sample bibliography file.

\section*{Acknowledgements}

% Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.

This research was funded by U.S. National Science Foundation (NSF) Grant ECCS 0941561 on
Cyberenabled Discovery and Innovation (CDI).

\section*{Author contributions statement}

% Must include all authors, identified by initials, for example:
% A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.

C.R. and H.L. conceived the experiments, C.R. conducted the experiments, and C.R. and H.L. analyzed
the results. Both authors reviewed the manuscript.

% \section*{Additional information}
%
% To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing interests} (mandatory statement). 
%
% The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on behalf of all authors of the paper. This statement must be included in the submitted article file.

% \begin{figure}[ht]
% \centering
% \includegraphics[width=\linewidth]{figures/stream}
% \caption{Legend (350 words max). Example legend text.}
% \label{fig:stream}
% \end{figure}

% \begin{table}[ht]
% \centering
% \begin{tabular}{|l|l|l|}
% \hline
% Condition & n & p \\
% \hline
% A & 5 & 0.1 \\
% \hline
% B & 10 & 0.01 \\
% \hline
% \end{tabular}
% \caption{\label{tab:example}Legend (350 words max). Example legend text.}
% \end{table}

% Figures and tables can be referenced in LaTeX using the ref command, e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.

\end{document}

### 2021-01-10 code/fnregister.m
Initial commit of in-progress NSR20 code

@@ -1,0 +1,237 @@
+%% Data Registration and Packaging Code for 3D Printed Wing Experiments
+% Charles Richter
+% Revised: July 10, 2011
+
+% function [] = fnregister(folderName,toggleRegisterPlots,toggleEndPlots,kBandwidth)
+clear,clc,close all,toggleRegisterPlots=1;toggleEndPlots=1;kBandwidth=7.5;
+##folderName = '\\zeus\shared\Charlie Richter\FUTEK Experiments\Experiment 7-12\Wing 100sp 50lc 10uc for Thesis Plots\Notch 1\';
+% folderName ='\\zeus\shared\Charlie Richter\FUTEK Experiments\Experiment 7-12\Wing 40sp 60lc 10uc\Notch 1\';
+folderName = '/home/crichter/ccsl/Wing 100sp 50lc 10uc/Wing 100sp 50lc 10uc/Notch 1/';
+
+pkg load signal
+
+%% Import data
+load ([folderName 'compdata.mat']);kinematics(isnan(kinematics)) = 0; %#ok<NODEF>
+load ([folderName 'forcewindows.mat']);fwin = forcewindows;
+load ([folderName 'massprops.mat'])
+load ([folderName 'constants.mat'])
+spanTotal = span+innerSpan;chordTotal = chordLower+chordUpper;
+fwin = fwin-(impulse1force-1); % shift the ranges to post-truncation indices
+
+%% Representative Force Plots for Thesis Showing Filtering and PSD
+% flapfreq = 3.0518; % for force window #5
+% window = fwin(5,1):fwin(5,2);
+% data = force;
+% 
+% % Filter force
+% cutoff = flapfreq*kBandwidth;width = 4;
+% h=fdesign.lowpass('Fp,Fst,Ap,Ast',cutoff,cutoff+width,.1,60,1000);
+% d=design(h,'equiripple'); %Lowpass FIR filter
+% dataf1 = filtfilt(d.Numerator,1,data); %zero-phase filtering
+% dataf = dataf1(window);
+% 
+% h2=fdesign.lowpass('Fp,Fst,Ap,Ast',50,54,.1,60,1000);
+% d2=design(h2,'equiripple'); %Lowpass FIR filter
+% dataf12 = filtfilt(d2.Numerator,1,data); %zero-phase filtering
+% dataf2 = dataf12(window);
+% 
+% % Compute Amplitude Spectrum
[…]

### 2021-01-10 code/force_simulation.m

@@ -1,0 +1,770 @@
+%% Load Data
+
+clear,clc,close all
+%----
+##folderName = '/Users/charlesrichter/Desktop/CR Shared Drive/';
+##spreadsheet = dlmread([folderName 'selecteddata_defcorr_all_70'],',',2,0);
+folderName = '/home/crichter/ccsl/';
+spreadsheet = dlmread('/home/crichter/ccsl/selecteddata_defcorr_all_70.csv',',',2,0);
+% ----
+%22,32/33,38,42,51,56,58,61,85,105
+% withheld = [22,32,38,42,51,56,58,61,85,105];
+% 32,51,58
+
+% Capture only the withheld experiments for testing
+exps = [8 26 33 37 39 48 52 53 95 103]; % 1,78 Removed
+EXPERIMENT_USED_TO_SHOWCASE_EQ4_COMPONENTS = 16; %%%%%%%%%%%
+
+
+% totalexps = unique(spreadsheet(:,1));
+% exps = totalexps(totalexps>0);
+
+% nExps = length(exps);
+% eqdataExps = cell(nExps,1);
+% whichExp = cell(nExps,1);
+% forceExps = cell(nExps,1);
+% for i = 1:nExps
+%     eqdataExps{i} = spreadsheet(spreadsheet(:,1)==exps(i),2:end);
+%     forceExps{i} = eqdataExps{i}(:,1);
+% end
+% eqdata = eqdataExps{1};
+
+% 19, 46 - use 41 instead of 46
+
+totalexps = unique(spreadsheet(:,1));
+exps = totalexps(totalexps>0);
+nExps = max(exps);
+eqdataExps = cell(nExps,1);
+for i = 1:length(exps)
+    whichExp = unique(spreadsheet(spreadsheet(:,1)==exps(i),1));
+    eqdataExps{whichExp} = spreadsheet(spreadsheet(:,1)==whichExp,2:end);
+end
+
+
+% figure(4)
+% for bla2 = 1:length(exps)
+%     disp(exps(bla2))
+%     plot(eqdataExps{exps(bla2)}(:,1));
+%     pause
+% end
+
+
+%%
+% use 52 53 33 48
+% use 33 39 53 103? 37?
+% toplot = [95 53 33 48 16]; % for exp 94 stuff
+% toplot = [39 48 52 53];
+% toplot = [8 26 33 37];
+toplot = [16 33 48 39 16];
+%16 39 48 33
+
[…]

### 2021-01-10 code/modeleval_complete_defcorr.m

@@ -1,0 +1,490 @@
+%% Model Evaluation
+clear,clc,close all
+
+pkg load optim
+pkg load signal
+
+##addpath("/home/crichter/ccsl/Code_CCSL_Drive/Code/Model Evaluation")
+addpath("/home/crichter/ccsl/NSR20_Richter/code/models")
+
+%% Load Data
+% folderName = '\\zeus\shared\Charlie Richter\FUTEK Experiments\Experiment 7-12\';
+% [spreadsheet,~,~] = xlsread([folderName 'selecteddata_defcorr_all_70.xls']);
+%----
+folderName = '/home/crichter/ccsl/';
+spreadsheet = dlmread('/home/crichter/ccsl/selecteddata_defcorr_all_70.csv',',',2,0);
+%----
+%% Experiments for Testing
+
+exps = [8 26 33 37 39 48 52 53 95 103]; % 1,78 removed
+% exps = [22,32,38,42,51,56,58,61,85,105]; % Make sure this is consistent with force_simulation.m
+
+% totalexps = unique(spreadsheet(:,1));
+% exps = totalexps(totalexps>0);
+
+nExps = length(exps);
+
+forcepredict = cell(1,nExps);
+forcepredicteq = cell(1,nExps);
+mae1 = nan(1,nExps);mae1eq = nan(1,nExps);
+rmd1 = nan(1,nExps);rmd1eq = nan(1,nExps);
+eqdata = cell(nExps,1);
+force = cell(nExps,1);
+
+% Initialize Coefficients - MEAN VALUES CALCULATED FROM *ALL* EXPS
+beta_jw = nan(nExps,3); % [Cl Crotlift Cm22] = [1.6864 1.4752 0.7536], STD = [0.2389 0.9920 0.4233]
+beta_rw_liftam = nan(nExps,3); % [Cl Camtrans Camrotddot]  = [1.8429 0.4382 -0.3152], STD = [0.2430 0.5158 1.0283]
+beta_trans_rot = nan(nExps,2); % [Cl Crotlift] = [1.7361 -0.0908], STD = [0.2453 0.8649]
+beta_trans_amrot = nan(nExps,2); % [Cl Camrot] = [1.7420 0.0755], STD = [0.2432 0.6126]
[…]

### 2021-01-10 code/models/fn_eq1.m

@@ -1,0 +1,30 @@
+function output = fn_eq1(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C1 = coeffs(1);
+C2 = coeffs(2);
+C3 = coeffs(3);
+C4 = coeffs(4);
+C5 = coeffs(5);
+output = C1.*span.*Ycm.*defdot.*defdot.*cos(def)...
+    + C2.*strokedot.*span.*Xcm.*def...
+    - C3.*strokedot.*span.*def.*devddot...
+    - C4.*strokedot.*span.*def.*defdot...
+    - C5.*span.*devddot;
+
+% Coeffs = [1.5723863e-8 1.2146853e-7 1.2146853e-7 1.2146853e-7 7.0700918e-7]
+% 1.5723863e-8*span*Ycm*defdot*defdot*cos(def) + 1.2146853e-7*strokedot*span*Xcm*def - 1.2146853e-7*strokedot*span*def*devddot - 1.2146853e-7*strokedot*span*def*defdot - 7.0700918e-7*span*devddot

### 2021-01-10 code/models/fn_eq2.m

@@ -1,0 +1,29 @@
+function output = fn_eq2(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C1 = coeffs(1);
+C2 = coeffs(2);
+C3 = coeffs(3);
+C4 = coeffs(4);
+% C5 = coeffs(5);
+output = C1.*span.*Ycm.*defdot.*defdot.*cos(def)...
+    + C2.*strokedot.*span.*Xcm.*def...
+    - C3.*strokedot.*span.*def.*defdot...
+    - C4.*span.*devddot;
+
+% Coeffs = [1.419276e-8 1.2280232e-7 1.2280232e-7 1.1480129e-6]
+% 1.419276e-8*span*Ycm*defdot*defdot*cos(def) + 1.2280232e-7*strokedot*span*Xcm*def - 1.2280232e-7*strokedot*span*def*defdot - 1.1480129e-6*span*devddot

### 2021-01-10 code/models/fn_eq3.m

@@ -1,0 +1,29 @@
+function output = fn_eq3(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3.*eqdata(:,17);innerspan = 1e-3.*eqdata(:,18);outerspan = 1e-3.*eqdata(:,19);
+% chord = 1e-3.*eqdata(:,22);lowerchord = 1e-3.*eqdata(:,20);upperchord =
+% 1e-3.*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C1 = coeffs(1);
+C2 = coeffs(2);
+C3 = coeffs(3);
+% C4 = coeffs(4);
+output = C1.*span.*Ycm.*defdot.*defdot.*cos(def - def.*def)...
+    + C2.*strokedot.*span.*Xcm.*def...
+    - C3.*span.*devddot;
+
+% Coeffs = [1.4180717e-8 1.2409225e-7 1.0660193e-6]
+% 1.4180717e-8.*span.*Ycm.*defdot.*defdot.*cos(def - def.*def) +
+% 1.2409225e-7.*strokedot.*span.*Xcm.*def - 1.0660193e-6.*span.*devddot

### 2021-01-10 code/models/fn_eq4.m

@@ -1,0 +1,26 @@
+function output = fn_eq4(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C1 = coeffs(1);
+C2 = coeffs(2);
+C3 = coeffs(3);
+% C4 = coeffs(4);
+output = C1.*span.*Ycm.*defdot.*defdot.*cos(def)...
+    + C2.*strokedot.*span.*Xcm.*def...
+    - C3.*span.*devddot;
+
+% Coeffs = [1.4526972e-8 1.2489714e-7 1.0872654e-6]

### 2021-01-10 code/models/fn_eq5.m

@@ -1,0 +1,26 @@
+function output = fn_eq5(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C1 = coeffs(1);
+C2 = coeffs(2);
+C3 = coeffs(3);
+% C4 = coeffs(4);
+output = C1.*strokedot.*span.*Xcm.*def...
+    - C2.*span.*lowerchord.*lowerchord.*defdot.*defdot...
+    - C3.*span.*devddot;
+
+% Coeffs = [1.2524723e-7 5.8678208e-11 1.1539232e-6]

### 2021-01-10 code/models/fn_eq6.m

@@ -1,0 +1,26 @@
+function output = fn_eq6(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C1 = coeffs(1);
+C2 = coeffs(2);
+C3 = coeffs(3);
+% C4 = coeffs(4);
+output = C1.*span.*Ycm.*defdot.*defdot...
+    + C2.*strokedot.*span.*Xcm.*def...
+    - C3.*span.*devddot;
+
+% Coeffs = [1.3552796e-8 1.2505514e-7 1.1357268e-6]

### 2021-01-10 code/models/fn_jw.m

@@ -1,0 +1,50 @@
+function output = fn_jw(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% ZJW Model
+
+if length(coeffs)==3
+    Ct = coeffs(1); % 1.2
+    Cr = coeffs(2); % pi;
+    Cam = coeffs(3);
+    m22 = Cam*pi*rho*(chord/2).^2;
+else
+    Ct = coeffs(1); % 1.2
+    Cr = coeffs(2); % pi;
+    m22 = pi*rho*(chord/2).^2;
+end
+    
[…]

### 2021-01-10 code/models/fn_planarinertial.m

@@ -1,0 +1,23 @@
+function output = fn_planarinertial(eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+output = inertial;

### 2021-01-10 code/models/fn_planarlift_planarinertial.m

@@ -1,0 +1,33 @@
+function output = fn_planarlift_planarinertial(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% Planar U-Squared Lift, Planar Inertial Forces - COMPLEXITY 47
+% Unpack Coefficients:
+Clmax = coeffs(1);
+
+% Compute Force
+beta = atan2(strokedot,0);
+alpha = beta-def;
+Cl = Clmax*sin(2*alpha);
+lift = .5*rho.*chord*(1/3).*(outerspan.^3-innerspan.^3).*Cl.*strokedot.^2.*sin(beta);
+output = lift+inertial;

### 2021-01-10 code/models/fn_planarlift_totalinertial.m

@@ -1,0 +1,33 @@
+function output = fn_planarlift_totalinertial(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% Planar U-Squared Lift, Planar Inertial Forces - COMPLEXITY 47
+% Unpack Coefficients:
+Clmax = coeffs(1);
+
+% Compute Force
+beta = atan2(strokedot,0);
+alpha = beta-def;
+Cl = Clmax*sin(2*alpha);
+lift = .5*rho.*chord*(1/3).*(outerspan.^3-innerspan.^3).*Cl.*strokedot.^2.*sin(beta);
+output = lift+inertial+inerttheta;

### 2021-01-10 code/models/fn_rw_liftam.m

@@ -1,0 +1,61 @@
+function output = fn_rw_liftam(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% RJW Lift (includes deviation) - NOT FITTING AM COEFFS HERE, BUT WOULD BE
+%% WORTH DOING SO
+% Unpack Coefficients:
+
+if length(coeffs)==3
+    Clmax = coeffs(1);
+    betaz = coeffs(2);
+    betazw = coeffs(3);
+    lambdaz = betaz*pi*rho*(chord/2).^2; % kg/m^3 * m^2
[…]

### 2021-01-10 code/models/fn_rw_liftonly.m

@@ -1,0 +1,40 @@
+function output = fn_rw_liftonly(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% RJW Lift (includes deviation)
+% Unpack Coefficients:
+Clmax = coeffs(1);
+
+% Compute Force
+wx = defdot-strokedot.*sin(dev);
+wxdot = defddot - (strokeddot.*sin(dev)+strokedot.*devdot.*cos(dev));
+wz = devdot.*cos(def) + strokedot.*sin(def).*cos(dev); % Vo = r_i*omega_z
+wy = -strokedot.*cos(def).*cos(dev) + devdot.*sin(def); % Wo = r_i*omega_y
[…]

### 2021-01-10 code/models/fn_totalinertial.m

@@ -1,0 +1,23 @@
+function output = fn_totalinertial(eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+output = inertial+inerttheta;

### 2021-01-10 code/models/fn_trans_amrot.m

@@ -1,0 +1,41 @@
+function output = fn_trans_amrot(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% Trans_AMRot: Complexity ZJW - 17
+Ct = coeffs(1); % 1.2
+% Cr = coeffs(2); % pi;
+Camrot = coeffs(2);
+m22 = Camrot*pi*rho*(chord/2).^2;
+
+% wx = defdot-strokedot.*sin(dev);
+wz = devdot.*cos(def) + strokedot.*sin(def).*cos(dev);
[…]

### 2021-01-10 code/models/fn_trans_amrotddot.m

@@ -1,0 +1,51 @@
+function output = fn_trans_amrotddot(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% RJW Lift (includes deviation) - NOT FITTING AM COEFFS HERE, BUT WOULD BE
+%% WORTH DOING SO
+% Unpack Coefficients:
+Clmax = coeffs(1);
+Camrotddot = coeffs(2);
+
+% Compute Force
+wx = defdot-strokedot.*sin(dev);
+wxdot = defddot - (strokeddot.*sin(dev)+strokedot.*devdot.*cos(dev));
[…]

### 2021-01-10 code/models/fn_trans_amtrans.m

@@ -1,0 +1,41 @@
+function output = fn_trans_amtrans(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% Trans_AMRot: Complexity ZJW - 15
+Ct = coeffs(1); % 1.2
+% Cr = coeffs(2); % pi;
+Camtrans = coeffs(2);
+m22 = Camtrans*pi*rho*(chord/2).^2;
+
+wx = defdot-strokedot.*sin(dev);
+wz = devdot.*cos(def) + strokedot.*sin(def).*cos(dev);
[…]

### 2021-01-10 code/models/fn_trans_amtrans_amrot.m

@@ -1,0 +1,45 @@
+function output = fn_trans_amtrans_amrot(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% Trans_AMRot: Complexity ZJW - 10
+Ct = coeffs(1); % 1.2
+% Cr = coeffs(2); % pi;
+Camtrans = coeffs(2);
+Camrot = coeffs(3);
+m22trans = Camtrans*pi*rho*(chord/2).^2;
+m22rot = Camrot*pi*rho*(chord/2).^2;
+
+wx = defdot-strokedot.*sin(dev);
[…]

### 2021-01-10 code/models/fn_trans_rot.m

@@ -1,0 +1,42 @@
+function output = fn_trans_rot(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Other Constants
+rho720 = 1.2159e+003; % 1.092*(1/1000)*(100/1)^3; % kg/m^3
+thickness = 9.6000e-005; % m
+rho = 1.2; % kg/m^3, density of air
+
+%% Inertial
+inertial = -(chord/2-upperchord).*chord*thickness.*span*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
+inerttheta = -.5*(innerspan+outerspan).*span.*chord*thickness*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
+
+%% Trans_Rot: Complexity ZJW - 12
+Ct = coeffs(1); % 1.2
+Cr = coeffs(2); % pi; 1.9254;
+% m22 = pi*rho*(chord/2).^2;
+
+% wx = defdot-strokedot.*sin(dev);
+wz = devdot.*cos(def) + strokedot.*sin(def).*cos(dev);
[…]

### 2021-01-10 code/fnregister.m
Update directory names

@@ -205,8 +205,8 @@ for p = 5%1:size(fwin,1)
 %     set(h1,'units','inches','position',[-14 1 width+2 height+2])
 %     set(h1,'papersize',[width+1.5 height+1],'paperpositionmode','auto')
 ##    saveFolder = '\\zeus\shared\Charlie Richter\Thesis\figures\';
-    saveFolder = '/home/crichter/ccsl/NSR20_Richter/paper/figures/';
-
+##    saveFolder = '/home/crichter/ccsl/NSR20_Richter/paper/figures/';
+      saveFolder = '/home/crichter/ccsl/RSI21_Richter/paper/figures/';
 %     ylabel('Measured Lift Force (N)')
 %     ylabel('Stroke Angle \phi (\circ)')
 %     ylabel('Deflection Angle \psi (\circ)')

### 2021-01-10 code/force_simulation.m

@@ -621,7 +621,8 @@ text(axwidth-.24,axheight-.1,'B','units','centimeters',...
 
 % set(gcf,'units','inches','papersize',[6 6],'paperpositionmode','auto')
 ##saveFolder = '/Users/charlesrichter/Desktop/PNAS11_Richter/figures/';
-saveFolder = '/home/crichter/ccsl/NSR20_Richter/paper/figures/';
+##saveFolder = '/home/crichter/ccsl/NSR20_Richter/paper/figures/';
+saveFolder = '/home/crichter/ccsl/RSI21_Richter/paper/figures/';
 saveas(gcf,[saveFolder,'eq_comparison_PNAS'],'epsc')
 end
 
@@ -764,7 +765,8 @@ end
 
 if bla == 4
 ##saveFolder = '/Users/charlesrichter/Desktop/PNAS11_Richter/figures/';
-saveFolder = '/home/crichter/ccsl/NSR20_Richter/paper/figures/';
+##saveFolder = '/home/crichter/ccsl/NSR20_Richter/paper/figures/';
+saveFolder = '/home/crichter/ccsl/RSI21_Richter/paper/figures/';
 saveas(gcf,[saveFolder,'eq_goodbad_PNAS'],'epsc')
 end
 end

### 2021-01-10 code/modeleval_complete_defcorr.m

@@ -5,7 +5,8 @@ pkg load optim
 pkg load signal
 
 ##addpath("/home/crichter/ccsl/Code_CCSL_Drive/Code/Model Evaluation")
-addpath("/home/crichter/ccsl/NSR20_Richter/code/models")
+##addpath("/home/crichter/ccsl/NSR20_Richter/code/models")
+addpath("/home/crichter/ccsl/RSI21_Richter/code/models")
 
 %% Load Data
 % folderName = '\\zeus\shared\Charlie Richter\FUTEK Experiments\Experiment 7-12\';
@@ -391,7 +392,8 @@ set(gcf,'units','centimeters','paperunits','centimeters','papersize',[width heig
 
 % saveFolder = '/Users/charlesrichter/Desktop/CR Shared Drive/Thesis/figures/';
 ##saveFolder = '/Users/charlesrichter/Desktop/PNAS11_Richter/figures/';
-saveFolder = '/home/crichter/ccsl/NSR20_Richter/paper/figures/';
+##saveFolder = '/home/crichter/ccsl/NSR20_Richter/paper/figures/';
+saveFolder = '/home/crichter/ccsl/RSI21_Richter/paper/figures/';
 saveas(gcf,[saveFolder,'pareto_mae_nofit_PNAS'],'epsc')
 % MARKERSIZES, FONTSIZES, DIMENSIONS, ETC ARE CHANGED FROM THE ORIGINAL
 % THESIS PLOT TO BE USED IN THE JOURNAL PUBLICATION.  REFER TO THE PLOTS

### 2021-01-10 code/modeleval_fit_coeffs.m
WIP jointly fit analytical model coeffients to the training experiments

@@ -1,0 +1,201 @@
+clear,clc,close all
+
+pkg load optim
+pkg load signal
+
+addpath("/home/crichter/ccsl/RSI21_Richter/code/models")
+
+%% Load Data
+folderName = '/home/crichter/ccsl/';
+spreadsheet = dlmread('/home/crichter/ccsl/selecteddata_defcorr_all_70.csv',',',2,0);
+
+%% Determine training, testing and other removed experiment IDs
+exps_all = unique(spreadsheet(:,1));
+exps_testing = [8 26 33 37 39 48 52 53 95 103]; % 1,78 removed
+exps_removed = [1, 78];
+exps_training = setdiff(exps_all, [exps_testing, exps_removed, 0]);
+
+% Mask off data to only include rows for which the experiment number (first column) is not in exps_testing, exps_removed, or 0.
+should_include = any(repmat(spreadsheet(:,1),1,size(exps_training)(1)) == repmat(exps_training',size(spreadsheet)(1),1),2);
+size(should_include)
+
+eqdata = spreadsheet(should_include,2:end);
+force = spreadsheet(should_include,2);
+
+##forcepredict = cell(1,nExps);
+##forcepredicteq = cell(1,nExps);
+##mae1 = nan(1,nExps);mae1eq = nan(1,nExps);
+##rmd1 = nan(1,nExps);rmd1eq = nan(1,nExps);
+##eqdata = cell(nExps,1);
+##force = cell(nExps,1);
+
+% Initialize Coefficients - MEAN VALUES CALCULATED FROM *ALL* EXPS
+##beta_jw = nan(nExps,3); % [Cl Crotlift Cm22] = [1.6864 1.4752 0.7536], STD = [0.2389 0.9920 0.4233]
+##beta_rw_liftam = nan(nExps,3); % [Cl Camtrans Camrotddot]  = [1.8429 0.4382 -0.3152], STD = [0.2430 0.5158 1.0283]
+##beta_trans_rot = nan(nExps,2); % [Cl Crotlift] = [1.7361 -0.0908], STD = [0.2453 0.8649]
[…]

### 2021-01-17 code/modeleval_rsi21.m
Initial commit of model evaluation code for RSI21 submission

@@ -1,0 +1,719 @@
+clear,clc,close all
+pkg load optim
+pkg load signal
+addpath("/home/crichter/ccsl/RSI21_Richter/code/models")
+
+%% Load Data
+spreadsheet = dlmread('/home/crichter/ccsl/selecteddata_defcorr_all_70.csv',',',2,0);
+
+%% Determine training, testing and other removed experiment IDs
+exps_all = unique(spreadsheet(:,1));
+exps_testing = [8 26 33 37 39 48 52 53 95 103]; % 1,78 removed
+exps_removed = [1, 78];
+exps_training = setdiff(exps_all, [exps_testing, exps_removed, 0]);
+
+##% Mask off data to only include rows for which the experiment number (first column) is not in exps_testing, exps_removed, or 0.
+##exps_training_mask = any(repmat(spreadsheet(:,1),1,size(exps_training)(1)) == repmat(exps_training',size(spreadsheet)(1),1),2);
+##eqdata = spreadsheet(exps_training_mask,2:end);
+##force = spreadsheet(exps_training_mask,2);
+
+%% Coefficients of Eureqa models (fit by Eureqa to training data)
+beta_eq1nf = [1.5723863e-8 1.2146853e-7 1.2146853e-7 1.2146853e-7 7.0700918e-7];
+beta_eq2nf = [1.419276e-8 1.2280232e-7 1.2280232e-7 1.1480129e-6];
+beta_eq3nf = [1.4180717e-8 1.2409225e-7 1.0660193e-6];
+beta_eq4nf = [1.4526972e-8 1.2489714e-7 1.0872654e-6];
+beta_eq5nf = [1.2524723e-7 5.8678208e-11 1.1539232e-6];
+beta_eq6nf = [1.3552796e-8 1.2505514e-7 1.1357268e-6];
+
+%% Specify initial values for parameter fitting
+beta0_plpi = 1.5;
+beta0_plti = 1.5;
+beta0_rw_liftonly = 1.5;
+beta0_rw_liftam = [1.5 1 1];
+beta0_jw = [1.5 pi 1];
+beta0_trans_rot = [1.5 2];
+beta0_trans_amrot = [1.5 1];
[…]

### 2021-01-17 code/modeleval_rsi21.m
Add equation sizes

@@ -17,6 +17,43 @@ exps_training = setdiff(exps_all, [exps_testing, exps_removed, 0]);
 ##eqdata = spreadsheet(exps_training_mask,2:end);
 ##force = spreadsheet(exps_training_mask,2);
 
+%% Equation sizes
+sz_plpi = 32;
+sz_plti = 46;
+sz_rw_liftonly = 96;
+sz_rw_liftam = 121;
+sz_jw = 130;
+sz_trans_rot = 118;
+sz_trans_amrot = 113;
+sz_trans_amtrans = 115;
+sz_trans_amtrans_amrot = 120;
+sz_trans_amrotddot = 106;
+sz_eq1 = 24;
+sz_eq2 = 19;
+sz_eq3 = 16;
+sz_eq4 = 14;
+sz_eq5 = 13;
+sz_eq6 = 12;
+
+% Planar Inertial -- fn_planarinertial -- COMPLEXITY 13
+% Total Inertial -- fn_totalinertial -- COMPLEXITY 27
+% Planar Lift Planar Inertial -- fn_planarlift_planarinertial -- COMPLEXITY 32
+% Planar Lift Total Inertial -- fn_planarlift_totalinertial -- COMPLEXITY 46
+% RW Lift Only Total Inertial -- fn_rw_liftonly -- COMPLEXITY 96
+% RW Lift and AM Total Inertial -- fn_rw_liftam -- COMPLEXITY 121
+% JW Lift and Total Inertial -- fn_jw -- COMPLEXITY 130
+% Analytical: fn_trans_rot -- fn_trans_rot -- COMPLEXITY 118
+% Analytical: fn_trans_amrot -- fn_trans_amrot -- COMPLEXITY 113
+% Analytical: fn_trans_amtrans -- fn_trans_amtrans -- COMPLEXITY 115
+% Analytical: fn_trans_amtrans_amrot -- fn_trans_amtrans_amrot -- COMPLEXITY 120
+% Analytical: fn_trans_amrotddot -- fn_trans_amrotddot -- COMPLEXITY 106
+% EQ 1 -- fn_eq1 -- COMPLEXITY 24
+% EQ 2 -- fn_eq2 -- COMPLEXITY 19
+% EQ 3 -- fn_eq3 -- COMPLEXITY 16
+% EQ 4 -- fn_eq4 -- COMPLEXITY 14
+% EQ 5 -- fn_eq5 -- COMPLEXITY 13
[…]

### 2021-01-17 code/modeleval_rsi21.m
Assign mean MAE results to variables

@@ -174,24 +174,24 @@ end
 
 format short e
 disp("Analytical models for publication");
-mean(nofit_mae_plpi,2)
-mean(nofit_mae_plti,2)
-mean(nofit_mae_rw_liftonly,2)
-mean(nofit_mae_rw_liftam,2)
-mean(nofit_mae_jw,2)
+mean_nofit_mae_plpi = mean(nofit_mae_plpi,2)
+mean_nofit_mae_plti = mean(nofit_mae_plti,2)
+mean_nofit_mae_rw_liftonly = mean(nofit_mae_rw_liftonly,2)
+mean_nofit_mae_rw_liftam = mean(nofit_mae_rw_liftam,2)
+mean_nofit_mae_jw = mean(nofit_mae_jw,2)
 disp("Other models");
-mean(nofit_mae_trans_rot,2)
-mean(nofit_mae_trans_amrot,2)
-mean(nofit_mae_trans_amtrans,2)
-mean(nofit_mae_trans_amtrans_amrot,2)
-mean(nofit_mae_trans_amrotddot,2)
+mean_nofit_mae_trans_rot = mean(nofit_mae_trans_rot,2)
+mean_nofit_mae_trans_amrot = mean(nofit_mae_trans_amrot,2)
+mean_nofit_mae_trans_amtrans = mean(nofit_mae_trans_amtrans,2)
+mean_nofit_mae_trans_amtrans_amrot = mean(nofit_mae_trans_amtrans_amrot,2)
+mean_nofit_mae_trans_amrotddot = mean(nofit_mae_trans_amrotddot,2)
 disp("Eureqa models");
-mean(nofit_mae_eq1,2)
-mean(nofit_mae_eq2,2)
-mean(nofit_mae_eq3,2)
-mean(nofit_mae_eq4,2)
-##mean(nofit_mae_eq5,2)
-mean(nofit_mae_eq6,2)
+mean_nofit_mae_eq1 = mean(nofit_mae_eq1,2)
+mean_nofit_mae_eq2 = mean(nofit_mae_eq2,2)
+mean_nofit_mae_eq3 = mean(nofit_mae_eq3,2)
+mean_nofit_mae_eq4 = mean(nofit_mae_eq4,2)
+##mean_nofit_mae_eq5 = mean(nofit_mae_eq5,2)
+mean_nofit_mae_eq6 = mean(nofit_mae_eq6,2)
 format
 
 %% Test all models on test data using coefficients fit to TEST data

### 2021-01-17 code/modeleval_rsi21.m
Initial implementation of results figure

@@ -194,6 +194,36 @@ mean_nofit_mae_eq4 = mean(nofit_mae_eq4,2)
 mean_nofit_mae_eq6 = mean(nofit_mae_eq6,2)
 format
 
+figure(1),clf,hold on
+set(gcf,'units','inches','outerposition',[0 0 5.8 3])
+largesize = 20;
+smallsize = 16;
+rotangle = 30;
+fsize = 6;
+
+plot(sz_plpi,mean_nofit_mae_plpi,'b.','Markersize',largesize);
+text(sz_plpi,mean_nofit_mae_plpi,'   U^2 Planar 1','rotation',rotangle,'fontsize',fsize);
+plot(sz_plti,mean_nofit_mae_plti,'g.','Markersize',largesize);
+text(sz_plti,mean_nofit_mae_plti,'   U^2 Planar 2','rotation',rotangle,'fontsize',fsize);
+plot(sz_rw_liftonly,mean_nofit_mae_rw_liftonly,'c.','Markersize',largesize);
+text(sz_rw_liftonly,mean_nofit_mae_rw_liftonly,'   U^2 Total','rotation',rotangle,'fontsize',fsize);
+plot(sz_rw_liftam,mean_nofit_mae_rw_liftam,'r.','Markersize',largesize);
+text(sz_rw_liftam,mean_nofit_mae_rw_liftam,'   Whitney-Wood','rotation',rotangle,'fontsize',fsize);
+plot(sz_jw,mean_nofit_mae_jw,'m.','Markersize',largesize);
+text(sz_jw,mean_nofit_mae_jw,'   Pesavento-Wang','rotation',rotangle,'fontsize',fsize);
+plot(sz_eq1,mean_nofit_mae_eq1,'k.','Markersize',smallsize);
+plot(sz_eq2,mean_nofit_mae_eq2,'k.','Markersize',smallsize);
+plot(sz_eq3,mean_nofit_mae_eq3,'k.','Markersize',smallsize);
+plot(sz_eq4,mean_nofit_mae_eq4,'k.','Markersize',smallsize);
+plot(sz_eq6,mean_nofit_mae_eq6,'k.','Markersize',smallsize);
+text(sz_eq1-6,mean_nofit_mae_eq1-1.5e-4,{'EQ';'Models'},'horizontalalignment','center','fontsize',fsize);
+
+ylabel('Mean Absolute Error (N)','fontsize',fsize)
+xlabel('Equation Size (number of operators)','fontsize',fsize)
+ylim([0 19e-4])
[…]

### 2021-01-17 code/modeleval_rsi21.m
Minor changes to figure generation

@@ -3,6 +3,8 @@ pkg load optim
 pkg load signal
 addpath("/home/crichter/ccsl/RSI21_Richter/code/models")
 
+savefolder = '/home/crichter/ccsl/RSI21_Richter/paper/figures/';
+
 %% Load Data
 spreadsheet = dlmread('/home/crichter/ccsl/selecteddata_defcorr_all_70.csv',',',2,0);
 
@@ -199,7 +201,7 @@ set(gcf,'units','inches','outerposition',[0 0 5.8 3])
 largesize = 20;
 smallsize = 16;
 rotangle = 30;
-fsize = 6;
+fsize = 12;
 
 plot(sz_plpi,mean_nofit_mae_plpi,'b.','Markersize',largesize);
 text(sz_plpi,mean_nofit_mae_plpi,'   U^2 Planar 1','rotation',rotangle,'fontsize',fsize);
@@ -223,6 +225,7 @@ xlabel('Equation Size (number of operators)','fontsize',fsize)
 ylim([0 19e-4])
 xlim([0 180])
 set(gca,'box','on');
+saveas(gcf,[savefolder, 'mae_nofit.eps'],'epsc');
 
 %% Test all models on test data using coefficients fit to TEST data
 disp("Fitting and testing on testing data...");

### 2021-01-17 code/modeleval_rsi21.m
Modify figure parameters

@@ -196,12 +196,17 @@ mean_nofit_mae_eq4 = mean(nofit_mae_eq4,2)
 mean_nofit_mae_eq6 = mean(nofit_mae_eq6,2)
 format
 
-figure(1),clf,hold on
-set(gcf,'units','inches','outerposition',[0 0 5.8 3])
-largesize = 20;
-smallsize = 16;
+f = figure('visible','off');clf,hold on
+largesize = 30;
+smallsize = 20;
 rotangle = 30;
 fsize = 12;
+##width=8.7;height=5;
+width=16;height=10;
+a = get(f,'currentaxes');
+set(a,'box','on','layer','top','fontunits','points','fontsize',fsize)
+set(f,'units','centimeters','paperunits','centimeters','papersize',[width height],...
+    'paperposition',[0 0 width height])
 
 plot(sz_plpi,mean_nofit_mae_plpi,'b.','Markersize',largesize);
 text(sz_plpi,mean_nofit_mae_plpi,'   U^2 Planar 1','rotation',rotangle,'fontsize',fsize);
@@ -218,14 +223,24 @@ plot(sz_eq2,mean_nofit_mae_eq2,'k.','Markersize',smallsize);
 plot(sz_eq3,mean_nofit_mae_eq3,'k.','Markersize',smallsize);
 plot(sz_eq4,mean_nofit_mae_eq4,'k.','Markersize',smallsize);
 plot(sz_eq6,mean_nofit_mae_eq6,'k.','Markersize',smallsize);
-text(sz_eq1-6,mean_nofit_mae_eq1-1.5e-4,{'EQ';'Models'},'horizontalalignment','center','fontsize',fsize);
-
-ylabel('Mean Absolute Error (N)','fontsize',fsize)
+text(sz_eq1-6,mean_nofit_mae_eq1-1.8e-4,{'EQ';'Models'},'horizontalalignment','center','fontsize',fsize);
+ylabel('Mean Absolute Error (10^-^3 N)','fontsize',fsize)
 xlabel('Equation Size (number of operators)','fontsize',fsize)
-ylim([0 19e-4])
-xlim([0 180])
-set(gca,'box','on');
-saveas(gcf,[savefolder, 'mae_nofit.eps'],'epsc');
+
+xmin = 0;
+xmax = 180;
+ymin = 0;
+ymax = 1.9e-3;
+xlim([xmin xmax])
+ylim([ymin ymax])
+dy = 2e-4;
[…]

### 2021-01-17 paper/main.tex
Comment out kinematics figures

@@
 high speed video and processing the measured data\cite{richter2012thesis}.
-Figure~\ref{fig:kinematics} provides sample force data and measured wing kinematics from one
-experiment.
+% Figure~\ref{fig:kinematics} provides sample force data and measured wing kinematics from one
+% experiment.
 
@@
 
-\begin{figure}
-\centering
-\begin{subfigure}[b]{.45\linewidth}
-\includegraphics[width=\linewidth]{figures/regforce1}
-\caption{Measured vertical force.}
-\label{fig:regforce}
-\end{subfigure}
-\begin{subfigure}[b]{.45\linewidth}
-\includegraphics[width=\linewidth]{figures/regstroke1}
-\caption{Wing stroke angle.}
-\label{fig:regstroke}
-\end{subfigure}
+% \begin{figure}
+% \centering
+% \begin{subfigure}[b]{.45\linewidth}
+% \includegraphics[width=\linewidth]{figures/regforce1}
+% \caption{Measured vertical force.}
+% \label{fig:regforce}
+% \end{subfigure}
+% \begin{subfigure}[b]{.45\linewidth}
+% \includegraphics[width=\linewidth]{figures/regstroke1}
+% \caption{Wing stroke angle.}
+% \label{fig:regstroke}
+% \end{subfigure}
 
-\begin{subfigure}[b]{.45\linewidth}
-\includegraphics[width=\linewidth]{figures/regdef1}
-\caption{Wing deflection angle.}
-\label{fig:regdef}
-\end{subfigure}
-\begin{subfigure}[b]{.45\linewidth}
-\includegraphics[width=\linewidth]{figures/regdev1}
-\caption{Wing deviation angle.}
-\label{fig:regdev}
-\end{subfigure}
-\caption{Measured \subref{fig:regforce} vertical force, \subref{fig:regstroke} wing stroke angle,
-  \subref{fig:regdef} deflection angle, \subref{fig:regdev} and deviation angle from one experiment,
-  registered and averaged over 18 flapping periods.}
-\label{fig:kinematics}
-\end{figure}
+% \begin{subfigure}[b]{.45\linewidth}
+% \includegraphics[width=\linewidth]{figures/regdef1}
+% \caption{Wing deflection angle.}
+% \label{fig:regdef}
+% \end{subfigure}
+% \begin{subfigure}[b]{.45\linewidth}
+% \includegraphics[width=\linewidth]{figures/regdev1}
+% \caption{Wing deviation angle.}
+% \label{fig:regdev}
+% \end{subfigure}
+% \caption{Measured \subref{fig:regforce} vertical force, \subref{fig:regstroke} wing stroke angle,
+%   \subref{fig:regdef} deflection angle, \subref{fig:regdev} and deviation angle from one experiment,
+%   registered and averaged over 18 flapping periods.}
+% \label{fig:kinematics}
+% \end{figure}
 

### 2021-01-17 code/modeleval_rsi21.m
Assign mean MAE results to variables (for fitted results)

@@ -300,24 +300,24 @@ end
 
 format short e
 disp("Analytical models for publication");
-mean(fit_mae_plpi,2)
-mean(fit_mae_plti,2)
-mean(fit_mae_rw_liftonly,2)
-mean(fit_mae_rw_liftam,2)
-mean(fit_mae_jw,2)
-disp("Other models");
-mean(fit_mae_trans_rot,2)
-mean(fit_mae_trans_amrot,2)
-mean(fit_mae_trans_amtrans,2)
-mean(fit_mae_trans_amtrans_amrot,2)
-mean(fit_mae_trans_amrotddot,2)
+mean_fit_mae_plpi = mean(fit_mae_plpi,2)
+mean_fit_mae_plti = mean(fit_mae_plti,2)
+mean_fit_mae_rw_liftonly = mean(fit_mae_rw_liftonly,2)
+mean_fit_mae_rw_liftam = mean(fit_mae_rw_liftam,2)
+mean_fit_mae_jw = mean(fit_mae_jw,2)
+##disp("Other models");
+##mean_fit_mae_trans_rot = mean(fit_mae_trans_rot,2)
+##mean_fit_mae_trans_amrot = mean(fit_mae_trans_amrot,2)
+##mean_fit_mae_trans_amtrans = mean(fit_mae_trans_amtrans,2)
+##mean_fit_mae_trans_amtrans_amrot = mean(fit_mae_trans_amtrans_amrot,2)
+##mean_fit_mae_trans_amrotddot = mean(fit_mae_trans_amrotddot,2)
 disp("Eureqa models");
-mean(fit_mae_eq1,2)
-mean(fit_mae_eq2,2)
-mean(fit_mae_eq3,2)
-mean(fit_mae_eq4,2)
-##mean(fit_mae_eq5,2)
-mean(fit_mae_eq6,2)
+mean_fit_mae_eq1 = mean(fit_mae_eq1,2)
+mean_fit_mae_eq2 = mean(fit_mae_eq2,2)
+mean_fit_mae_eq3 = mean(fit_mae_eq3,2)
+mean_fit_mae_eq4 = mean(fit_mae_eq4,2)
+##mean_fit_mae_eq5 = mean(fit_mae_eq5,2)
+mean_fit_mae_eq6 = mean(fit_mae_eq6,2)
 format
 
 return

### 2021-01-17 code/modeleval_rsi21.m
Comment out calculations for unused models

@@ -88,18 +88,18 @@ for i = 1:length(exps_training)
   beta_rw_liftonly(i,:) = nlinfit(eqdata,force,@fn_rw_liftonly,beta0_rw_liftonly);
   beta_rw_liftam(i,:) = nlinfit(eqdata,force,@fn_rw_liftam,beta0_rw_liftam);
   beta_jw(i,:) = nlinfit(eqdata,force,@fn_jw,beta0_jw);
-  beta_trans_rot(i,:) = nlinfit(eqdata,force,@fn_trans_rot,beta0_trans_rot);
-  beta_trans_amrot(i,:) = nlinfit(eqdata,force,@fn_trans_amrot,beta0_trans_amrot);
-  beta_trans_amtrans(i,:) = nlinfit(eqdata,force,@fn_trans_amtrans,beta0_trans_amtrans);
-  beta_trans_amtrans_amrot(i,:) = nlinfit(eqdata,force,@fn_trans_amtrans_amrot,beta0_trans_amtrans_amrot);
-  beta_trans_amrotddot(i,:) = nlinfit(eqdata,force,@fn_trans_amrotddot,beta0_trans_amrotddot);
+##  beta_trans_rot(i,:) = nlinfit(eqdata,force,@fn_trans_rot,beta0_trans_rot);
+##  beta_trans_amrot(i,:) = nlinfit(eqdata,force,@fn_trans_amrot,beta0_trans_amrot);
+##  beta_trans_amtrans(i,:) = nlinfit(eqdata,force,@fn_trans_amtrans,beta0_trans_amtrans);
+##  beta_trans_amtrans_amrot(i,:) = nlinfit(eqdata,force,@fn_trans_amtrans_amrot,beta0_trans_amtrans_amrot);
+##  beta_trans_amrotddot(i,:) = nlinfit(eqdata,force,@fn_trans_amrotddot,beta0_trans_amrotddot);
   
-  beta_eq1(i,:) = nlinfit(eqdata,force,@fn_eq1,beta_eq1nf);
-  beta_eq2(i,:) = nlinfit(eqdata,force,@fn_eq2,beta_eq2nf);
-  beta_eq3(i,:) = nlinfit(eqdata,force,@fn_eq3,beta_eq3nf);
-  beta_eq4(i,:) = nlinfit(eqdata,force,@fn_eq4,beta_eq4nf);
-  beta_eq5(i,:) = nlinfit(eqdata,force,@fn_eq5,beta_eq5nf);
[…]

### 2021-01-17 code/modeleval_rsi21.m
Minor changes to figure parameters

@@ -196,16 +196,14 @@ mean_nofit_mae_eq4 = mean(nofit_mae_eq4,2)
 mean_nofit_mae_eq6 = mean(nofit_mae_eq6,2)
 format
 
-f = figure('visible','off');clf,hold on
+f1 = figure('visible','off');clf,hold on
 largesize = 30;
 smallsize = 20;
 rotangle = 30;
 fsize = 12;
-##width=8.7;height=5;
 width=16;height=10;
-a = get(f,'currentaxes');
-set(a,'box','on','layer','top','fontunits','points','fontsize',fsize)
-set(f,'units','centimeters','paperunits','centimeters','papersize',[width height],...
+set(get(f1,'currentaxes'),'box','on','layer','top','fontunits','points','fontsize',fsize)
+set(f1,'units','centimeters','paperunits','centimeters','papersize',[width height],...
     'paperposition',[0 0 width height])
 
 plot(sz_plpi,mean_nofit_mae_plpi,'b.','Markersize',largesize);
@@ -240,7 +238,7 @@ yticklabels(yt*1e3);
 dx = 20;
 xt = xmin:dx:xmax;
 xticks(xt)
-saveas(f,[savefolder, 'mae_nofit.eps'],'epsc');
+saveas(f1,[savefolder, 'mae_nofit.eps'],'epsc');
 
 %% Test all models on test data using coefficients fit to TEST data
 disp("Fitting and testing on testing data...");

### 2021-01-17 code/modeleval_rsi21.m
Add plot for fit results

@@ -196,6 +196,7 @@ mean_nofit_mae_eq4 = mean(nofit_mae_eq4,2)
 mean_nofit_mae_eq6 = mean(nofit_mae_eq6,2)
 format
 
+%% Plot nofit results
 f1 = figure('visible','off');clf,hold on
 largesize = 30;
 smallsize = 20;
@@ -317,6 +318,51 @@ mean_fit_mae_eq4 = mean(fit_mae_eq4,2)
 ##mean_fit_mae_eq5 = mean(fit_mae_eq5,2)
 mean_fit_mae_eq6 = mean(fit_mae_eq6,2)
 format
+
+%% Plot fit results
+f2 = figure('visible','off');clf,hold on
+largesize = 30;
+smallsize = 20;
+rotangle = 30;
+fsize = 12;
+width=16;height=10;
+set(get(f2,'currentaxes'),'box','on','layer','top','fontunits','points','fontsize',fsize)
+set(f2,'units','centimeters','paperunits','centimeters','papersize',[width height],...
+    'paperposition',[0 0 width height])
+
+plot(sz_plpi,mean_fit_mae_plpi,'b.','Markersize',largesize);
+text(sz_plpi,mean_fit_mae_plpi,'   U^2 Planar 1','rotation',rotangle,'fontsize',fsize);
+plot(sz_plti,mean_fit_mae_plti,'g.','Markersize',largesize);
+text(sz_plti,mean_fit_mae_plti,'   U^2 Planar 2','rotation',rotangle,'fontsize',fsize);
+plot(sz_rw_liftonly,mean_fit_mae_rw_liftonly,'c.','Markersize',largesize);
+text(sz_rw_liftonly,mean_fit_mae_rw_liftonly,'   U^2 Total','rotation',rotangle,'fontsize',fsize);
+plot(sz_rw_liftam,mean_fit_mae_rw_liftam,'r.','Markersize',largesize);
+text(sz_rw_liftam,mean_fit_mae_rw_liftam,'   Whitney-Wood','rotation',rotangle,'fontsize',fsize);
+plot(sz_jw,mean_fit_mae_jw,'m.','Markersize',largesize);
+text(sz_jw,mean_fit_mae_jw,'   Pesavento-Wang','rotation',rotangle,'fontsize',fsize);
[…]

### 2021-01-17 code/forcesim_rsi21.m
Initial commit of force sim code

@@ -1,0 +1,282 @@
+clear,clc,close all
+folderName = '/home/crichter/ccsl/';
+spreadsheet = dlmread('/home/crichter/ccsl/selecteddata_defcorr_all_70.csv',',',2,0);
+
+% Capture only the withheld experiments for testing
+exps = [8 26 33 37 39 48 52 53 95 103]; % 1,78 Removed
+EXPERIMENT_USED_TO_SHOWCASE_EQ4_COMPONENTS = 16; %%%%%%%%%%%
+
+totalexps = unique(spreadsheet(:,1));
+exps = totalexps(totalexps>0);
+nExps = max(exps);
+eqdataExps = cell(nExps,1);
+for i = 1:length(exps)
+    whichExp = unique(spreadsheet(spreadsheet(:,1)==exps(i),1));
+    eqdataExps{whichExp} = spreadsheet(spreadsheet(:,1)==whichExp,2:end);
+end
+
+toplot = [16 33 48 39 16];
+
+for bla = 1:length(toplot)
+eqdata = eqdataExps{toplot(bla)};
+
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+thickness = .000096; % m
+rho720 = 1215.92; % kg/m^3
+rho = 1.2; % kg/m^3, density of air
+a = chord/2; % m
+yh = a-upperchord; % m
[…]

### 2021-01-17 code/models/fn_eq4_term_1.m

@@ -1,0 +1,21 @@
+function output = fn_eq4_term_1(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C1 = coeffs(1);
+output = C1.*span.*Ycm.*defdot.*defdot.*cos(def);
+    
+% Coeffs = [1.4526972e-8]

### 2021-01-17 code/models/fn_eq4_term_2.m

@@ -1,0 +1,21 @@
+function output = fn_eq4_term_2(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C2 = coeffs(1);
+output = C2.*strokedot.*span.*Xcm.*def;
+
+% Coeffs = [1.2489714e-7]

### 2021-01-17 code/models/fn_eq4_term_3.m

@@ -1,0 +1,21 @@
+function output = fn_eq4_term_3(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C3 = coeffs(1);
+output = -C3.*span.*devddot;
+
+% Coeffs = [1.0872654e-6]

### 2021-01-17 code/models/fn_eqbad.m

@@ -1,0 +1,32 @@
+function output = fn_eqbad(coeffs,eqdata)
+%% Unpack Data
+force = eqdata(:,1);
+stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
+def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
+chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
+Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
+Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
+Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+
+%% Compute Output
+C1 = coeffs(1);
+C2 = coeffs(2);
+C3 = coeffs(3);
+C4 = coeffs(4);
+C5 = coeffs(5);
+C6 = coeffs(6);
+C7 = coeffs(7);
+C8 = coeffs(8);
+C9 = coeffs(9);
+
+output = (14.9488 + 13.5087*def.*def - 1.38486.*def)./...
+  (407.258 + devddot + 17.8828.*devdot +...
+  cos(2.69241 - 42.9085./(13.5087.*def.*def) - 163.587.*def)) - 0.0381286;
+
+% Coeffs = [14.9488, 13.5087, 1.38486, 407.258, 17.8828, 2.69241, 42.9085, 13.5087, 163.587, 0.0381286]

### 2021-01-18 code/models/fn_rw_liftonly.m
Provide individual terms as additional outputs from fn_rw_liftonly

@@ -1,4 +1,4 @@
-function output = fn_rw_liftonly(coeffs,eqdata)
+function [output,Lrjw,inertial,inerttheta] = fn_rw_liftonly(coeffs,eqdata)
 %% Unpack Data
 force = eqdata(:,1);
 stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);

### 2021-01-18 code/forcesim_rsi21.m
Update number of sig figs in material density to be consistent with other functions

@@ -34,7 +34,8 @@ Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
 Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
 
 thickness = .000096; % m
-rho720 = 1215.92; % kg/m^3
+rho720 = 1215.9; % kg/m^3
+##rho720 = 1215.92; % kg/m^3 <<-- Changed to 1215.9 to be consistent with other fn's.
 rho = 1.2; % kg/m^3, density of air
 a = chord/2; % m
 yh = a-upperchord; % m

### 2021-01-18 code/forcesim_rsi21.m
Comment out in-place calculations in favor of the ones in defined function files

@@ -22,53 +22,63 @@ eqdata = eqdataExps{toplot(bla)};
 
 %% Unpack Data
 force = eqdata(:,1);
-stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
-def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
-dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
-% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
-span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
-chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
-Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
-Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
-Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
-Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
+##stroke = eqdata(:,2);
+strokedot = eqdata(:,3);
+##strokeddot = eqdata(:,4);
+##def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
+##dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
+##% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
+##span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
+##chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
+##Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
+##Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
[…]

### 2021-01-18 code/forcesim_rsi21.m
Delete unused commented code

@@ -22,90 +22,19 @@ eqdata = eqdataExps{toplot(bla)};
 
 %% Unpack Data
 force = eqdata(:,1);
-##stroke = eqdata(:,2);
 strokedot = eqdata(:,3);
-##strokeddot = eqdata(:,4);
-##def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
-##dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
-##% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
-##span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
-##chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
-##Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
-##Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
-##Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
-##Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
-
-##thickness = .000096; % m
-##rho720 = 1215.9; % kg/m^3
-####rho720 = 1215.92; % kg/m^3 <<-- Changed to 1215.9 to be consistent with other fn's.
-##rho = 1.2; % kg/m^3, density of air
-##a = chord/2; % m
-##yh = a-upperchord; % m
-##xh = (innerspan+outerspan)/2; % m
-##lambdaz = pi*rho*a.^2; % kg/m^3 * m^2
-##lambdazw = -pi*rho*a.^2.*yh; % kg/m^3 * m^2 * m
-
-##%% Inertial Reaction
-##inertial = -yh.*chord.*thickness.*span.*rho720.*(defddot.*sin(def) + defdot.^2.*cos(def));
-##inertialdev = -xh.*span.*chord.*thickness.*rho720.*(devddot.*cos(dev) - devdot.^2.*sin(dev));
-##
-##% Also calculated with full integration written out:
[…]

### 2021-01-18 code/models/fn_eq4.m
Provide individual terms as additional outputs from fn_eq4

@@ -1,4 +1,4 @@
-function output = fn_eq4(coeffs,eqdata)
+function [output,term1,term2,term3] = fn_eq4(coeffs,eqdata)
 %% Unpack Data
 force = eqdata(:,1);
 stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
@@ -18,9 +18,14 @@ Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
 C1 = coeffs(1);
 C2 = coeffs(2);
 C3 = coeffs(3);
-% C4 = coeffs(4);
-output = C1.*span.*Ycm.*defdot.*defdot.*cos(def)...
-    + C2.*strokedot.*span.*Xcm.*def...
-    - C3.*span.*devddot;
+
+term1 = C1.*span.*Ycm.*defdot.*defdot.*cos(def);
+term2 = C2.*strokedot.*span.*Xcm.*def;
+term3 = -C3.*span.*devddot;
+output = term1 + term2 + term3;
+
+% output = C1.*span.*Ycm.*defdot.*defdot.*cos(def)...
+%     + C2.*strokedot.*span.*Xcm.*def...
+%     - C3.*span.*devddot;
 
 % Coeffs = [1.4526972e-8 1.2489714e-7 1.0872654e-6]

### 2021-01-18 code/forcesim_rsi21.m
Use individual term outputs from fn_eq4

@@ -30,9 +30,7 @@ Clmax = 1.9;
 
 %% EQ4 (individual terms) Calculations
 beta_eq4 = [1.4526972e-8 1.2489714e-7 1.0872654e-6];
-eq1 = fn_eq4_term_1(beta_eq4(1),eqdata);
-eq2 = fn_eq4_term_2(beta_eq4(2),eqdata);
-eq3 = fn_eq4_term_3(beta_eq4(3),eqdata);
+[~,eq1,eq2,eq3] = fn_eq4(beta_eq4,eqdata);
 
 % Overfit equation
 beta_eqbad = [14.9488, 13.5087, 1.38486, 407.258, 17.8828, 2.69241, 42.9085, 13.5087, 163.587, 0.0381286];

### 2021-01-18 code/models/fn_eq4_term_1.m
Remove individual functions for calculating the terms of eq4

@@ -1,21 +1,0 @@
-function output = fn_eq4_term_1(coeffs,eqdata)
-%% Unpack Data
-force = eqdata(:,1);
-stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
-def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
-dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
-% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
-% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
-% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
-span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
-chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
-Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
-Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
-Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
-Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
-
-%% Compute Output
-C1 = coeffs(1);
-output = C1.*span.*Ycm.*defdot.*defdot.*cos(def);
-    
-% Coeffs = [1.4526972e-8]

### 2021-01-18 code/models/fn_eq4_term_2.m

@@ -1,21 +1,0 @@
-function output = fn_eq4_term_2(coeffs,eqdata)
-%% Unpack Data
-force = eqdata(:,1);
-stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
-def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
-dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
-% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
-% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
-% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
-span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
-chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
-Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
-Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
-Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
-Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
-
-%% Compute Output
-C2 = coeffs(1);
-output = C2.*strokedot.*span.*Xcm.*def;
-
-% Coeffs = [1.2489714e-7]

### 2021-01-18 code/models/fn_eq4_term_3.m

@@ -1,21 +1,0 @@
-function output = fn_eq4_term_3(coeffs,eqdata)
-%% Unpack Data
-force = eqdata(:,1);
-stroke = eqdata(:,2);strokedot = eqdata(:,3);strokeddot = eqdata(:,4);
-def = eqdata(:,5);defdot = eqdata(:,6);defddot = eqdata(:,7);
-dev = eqdata(:,11);devdot = eqdata(:,12);devddot = eqdata(:,13);
-% dev = zeros(size(dev));devdot = zeros(size(devdot));devddot = zeros(size(devddot));
-% span = 1e-3*eqdata(:,17);innerspan = 1e-3*eqdata(:,18);outerspan = 1e-3*eqdata(:,19);
-% chord = 1e-3*eqdata(:,22);lowerchord = 1e-3*eqdata(:,20);upperchord = 1e-3*eqdata(:,21);
-span = eqdata(:,17);innerspan = eqdata(:,18);outerspan = eqdata(:,19);
-chord = eqdata(:,22);lowerchord = eqdata(:,20);upperchord = eqdata(:,21);
-Xcm	= eqdata(:,23);Ycm	= eqdata(:,24);Zcm= eqdata(:,25);
-Ixx	= eqdata(:,26);Ixy	= eqdata(:,27);Ixz= eqdata(:,28);
-Iyx	= eqdata(:,29);Iyy	= eqdata(:,30);Iyz= eqdata(:,31);
-Izx	= eqdata(:,32);Izy	= eqdata(:,33);Izz= eqdata(:,34);
-
-%% Compute Output
-C3 = coeffs(1);
-output = -C3.*span.*devddot;
-
-% Coeffs = [1.0872654e-6]

### 2021-01-18 code/forcesim_rsi21.m
Cleanup

@@ -3,22 +3,11 @@ folderName = '/home/crichter/ccsl/';
 spreadsheet = dlmread('/home/crichter/ccsl/selecteddata_defcorr_all_70.csv',',',2,0);
 
 % Capture only the withheld experiments for testing
-exps = [8 26 33 37 39 48 52 53 95 103]; % 1,78 Removed
-EXPERIMENT_USED_TO_SHOWCASE_EQ4_COMPONENTS = 16; %%%%%%%%%%%
-
-totalexps = unique(spreadsheet(:,1));
-exps = totalexps(totalexps>0);
-nExps = max(exps);
-eqdataExps = cell(nExps,1);
-for i = 1:length(exps)
-    whichExp = unique(spreadsheet(spreadsheet(:,1)==exps(i),1));
-    eqdataExps{whichExp} = spreadsheet(spreadsheet(:,1)==whichExp,2:end);
-end
 
 toplot = [16 33 48 39 16];
 
-for bla = 1:length(toplot)
-eqdata = eqdataExps{toplot(bla)};
+for i_exp = 1:length(toplot)
+eqdata = spreadsheet(spreadsheet(:,1)==toplot(i_exp),2:end);
 
 %% Unpack Data
 force = eqdata(:,1);
@@ -43,7 +32,7 @@ domain = 0:length(force)-1;
 axbuf = 1.2;
 shadecolor = .9*[1 1 1];
 
-if bla == 5
+if i_exp == 5
     
 fontsz = 6;
 widthcm = 8.7;
@@ -130,7 +119,7 @@ saveFolder = '/home/crichter/ccsl/RSI21_Richter/paper/figures/';
 end
 
 %% Plot for PNAS
-if bla < 5
+if i_exp < 5
 txbuf = .2;
 fontsz = 6;
 widthcm2 = 17.8;
@@ -182,7 +171,7 @@ set(gcf,'units','centimeters','paperunits','centimeters','papersize',[widthcm2 h
     'paperposition',[0 0 widthcm2 heightcm2])
 set(gcf,'defaultlinelinewidth',.5,'visible','on')
 
-axes('units','centimeters','position',[plotgrid{bla} axwidth2 axheight2]),hold on
+axes('units','centimeters','position',[plotgrid{i_exp} axwidth2 axheight2]),hold on
 
 cross1 = find([0;diff(sign(strokedot))]);
 if cross1(1)>50
[…]

### 2021-01-18 code/forcesim_rsi21.m
Format and save figures from forcesim_rsi21.m

@@ -1,13 +1,11 @@
 clear,clc,close all
 folderName = '/home/crichter/ccsl/';
 spreadsheet = dlmread('/home/crichter/ccsl/selecteddata_defcorr_all_70.csv',',',2,0);
+saveFolder = '/home/crichter/ccsl/RSI21_Richter/paper/figures/';
 
-% Capture only the withheld experiments for testing
-
-toplot = [16 33 48 39 16];
-
-for i_exp = 1:length(toplot)
-eqdata = spreadsheet(spreadsheet(:,1)==toplot(i_exp),2:end);
+%% Compare terms of EQ4 with analytical model components
+n_exp = 16;
+eqdata = spreadsheet(spreadsheet(:,1)==n_exp,2:end);
 
 %% Unpack Data
 force = eqdata(:,1);
@@ -21,157 +19,88 @@ Clmax = 1.9;
 beta_eq4 = [1.4526972e-8 1.2489714e-7 1.0872654e-6];
 [~,eq1,eq2,eq3] = fn_eq4(beta_eq4,eqdata);
 
+%% Plotting Constants
+axbuf = 1.2;
+shadecolor = .9*[1 1 1];
+fontsz = 12;
+plotfactor = 1000; % Multiply force by 1000 then explain that all forces are "e-3"
+
+cross1 = find([0;diff(sign(strokedot))]);
+normstroke = [1;cross1;length(force)];
+xticklab = {'0','.5','1.0','1.5','2'};
+
+width = 15;
+height = 7;
+
+f1 = figure('visible','off');hold on
+set(f1, 'defaultlinelinewidth',1)
+fill([0 cross1(1) cross1(1) 0],10000*[-1 -1 1 1],shadecolor,'edgecolor',shadecolor)
+fill([cross1(2) cross1(3) cross1(3) cross1(2)],10000*[-1 -1 1 1],shadecolor,'edgecolor',shadecolor)
+plot(force*plotfactor,'k')
+plot(eq2*plotfactor,'b')
+plot(Lrjw*plotfactor,'r')
+
+ylimit = axbuf*[min([force*plotfactor;(eq1+eq3)*plotfactor]) max([force*plotfactor;(eq1+eq3)*plotfactor])];
+ylim(ylimit)
+xlim([1 length(force)])
+set(f1,'units','centimeters','paperunits','centimeters','papersize',[width height],...
+    'paperposition',[0 0 width height])
[…]

### 2021-01-18 paper/main.tex
Replace eq_comparison_PNAS with individual subfigures generated by forcesim_rsi21.m

@@
 \centering
-\includegraphics[width=.48\textwidth]{figures/eq_comparison_PNAS}
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/eq_comparison_lift}
+\caption{\label{fig:eqcomparison_lift}}
+\end{subfigure}
+% \qquad
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/eq_comparison_inertial}
+\caption{\label{fig:eqcomparison_inertial}}
+\end{subfigure}
 \caption{Comparison between terms of the analytical expressions (red) and data-driven models (blue).
-  (A) Primary translational aerodynamic lift proportional to the square of the wing velocity. (B)
-  Inertial forces related to the pendulum motion of the wing and deviation out of the stroke plane.
-  The modeling process automatically decoupled aerodynamic and inertial effects with no prior
-  knowledge of the system.  The total measured force is shown for reference (black).}
+  \subref{fig:eqcomparison_lift} Primary translational aerodynamic lift proportional to the square
+  of the wing velocity. \subref{fig:eqcomparison_inertial} Inertial forces related to the pendulum
+  motion of the wing and deviation out of the stroke plane. The modeling process automatically
+  decoupled aerodynamic and inertial effects with no prior knowledge of the system. The total
+  measured force is shown for reference (black).}
 \label{fig:eqcomparison}

### 2021-01-18 paper/main.tex
Replace pareto_mae_nofit_PNAS with mae_nofit generated by modeleval_rsi21.m
Replace eq_goodbad_PNAS with individual subfigures generated by forcesim_rsi21.m

@@
 \centering
-\includegraphics[width=\textwidth]{figures/pareto_mae_nofit_PNAS}
+\includegraphics[width=\textwidth]{figures/mae_nofit}
 \caption{\label{fig:pareto_mae}}
@@
 \centering
-\includegraphics[width=\textwidth]{figures/eq_goodbad_PNAS}
-\caption{Comparison of force predictions between $EQ_4$ (blue) and Equation~\eqref{eq:badeq} (red).
-  $EQ_4$ was trained on 58 distinct experiments simultaneously, including the data shown in (A),
-  whereas Equation~\eqref{eq:badeq} was trained \emph{only} on the data shown in (A).  Panel (A)
-  shows that both models perform well on this data within their training sets.  However, when
-  presented with data outside their training set (B,C,D), only $EQ_4$ is able to generalize and
-  provide accurate predictions, while Equation~\eqref{eq:badeq} fails to do so.}
+\begin{tabular}{cc}
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/eq_goodbad_1}
+\caption{\label{fig:eq_goodbad_1}}
+\end{subfigure} &
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/eq_goodbad_2}
+\caption{\label{fig:eq_goodbad_2}}
+\end{subfigure} \\
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/eq_goodbad_3}
+\caption{\label{fig:eq_goodbad_3}}
+\end{subfigure} &
+\begin{subfigure}{0.48\textwidth}
+\centering
+\includegraphics[width=\textwidth]{figures/eq_goodbad_4}
+\caption{\label{fig:eq_goodbad_4}}
+\end{subfigure}
+\end{tabular}
+\caption{Comparison of force predictions between $EQ_4$ (blue) and
+Equation~\eqref{eq:badeq} (red). $EQ_4$ was trained on 58 distinct experiments
+simultaneously, including the data shown in~\subref{fig:eq_goodbad_1}, whereas
+Equation~\eqref{eq:badeq} was trained \emph{only} on the data shown
+in~\subref{fig:eq_goodbad_1}.  \subref{fig:eq_goodbad_1} shows that both models
+perform well on this data within their training sets.  However, when presented
+with test data \subref{fig:eq_goodbad_2}, \subref{fig:eq_goodbad_3} and
+\subref{fig:eq_goodbad_4} not contained within either model's training set, only $EQ_4$ is able to
+generalize and provide accurate predictions, while Equation~\eqref{eq:badeq}
+fails to do so.}
 \label{fig:eq_goodbad}

### 2021-01-18 code/forcesim_rsi21.m
Change figure size

@@ -93,7 +93,7 @@ shadecolor = .9*[1 1 1];
 
 fontsz = 12;
 widthcm2 = 15;
-heightcm2 = 9;
+heightcm2 = 7;
 plotfactor = 1000; % Multiply force by 1000 then explain that all forces are "e-3"
 xticklab = {'0','.5','1.0','1.5','2'};
         

### 2021-01-18 paper/main.tex
Update figure references

@@
 \subsection{Performance of models trained on one experiment versus many experiments}
-Figure~\ref{fig:eq_goodbad}(A) illustrates the predictive performance of two models on data within their training sets, while Figure~\ref{fig:eq_goodbad}(B,C,D) illustrate the performance of those same two models on data outside their training sets.
-The differences between these two models result from the fact that one of them, $EQ_4$ from Table~\ref{table:eqmodels} (drawn in blue), was trained on 58 separate experiments simultaneously, whereas the other one, shown below (drawn in red), is one of many models trained only on the single experiment shown in Figure~\ref{fig:eq_goodbad}(A):
+Figure~\ref{fig:eq_goodbad_1} illustrates the predictive performance of two models on data within
+their training sets, while Figures~\ref{fig:eq_goodbad_2},~\ref{fig:eq_goodbad_3},
+and~\ref{fig:eq_goodbad_4} illustrate the performance of those same two models on data outside
+their training sets. The differences between these two models result from the fact that one of them,
+$EQ_4$ from Table~\ref{table:eqmodels} (drawn in blue), was trained on 58 separate experiments
+simultaneously, whereas the other one, shown below (drawn in red), is one of many models trained
+only on the single experiment shown in Figure~\ref{fig:eq_goodbad_1}:
 \begin{align}

### 2021-01-18 paper/main.tex
Add subfigure captions
Reformat paper to use basic article class, and associated changes

@@
-\documentclass[fleqn,10pt]{wlscirep}
+\documentclass{article}
+\usepackage{natbib}
+\usepackage{hyperref}
+\usepackage{authblk}
+\usepackage{fullpage}
+
+% \documentclass[fleqn,10pt]{wlscirep}
 \usepackage[utf8]{inputenc}
@@
 
+% Keywords command
+\providecommand{\keywords}[1]
+{
+  \vspace{5mm}
+  \small
+  \noindent
+  \textbf{Keywords:} #1
+  \normalsize
+}
+
 \title{Automated data-driven symbolic modeling of the lift forces acting on flapping wings}
+\date{}
 
@@
 \author[2,3]{Hod Lipson}
-% \author[1,2,+]{Christine Author}
-% \author[2,+]{Derek Author}
-\affil[1]{Cornell University, Sibley School of Mechanical and Aerospace
-Engineering, Ithaca, 14850, United States}
-\affil[2]{Columbia University, Department of Mechanical Engineering, New York,
-10027, United States}
-\affil[3]{Columbia University, Data Science Institute, New York, 10027, United
-States}
-
+\affil[1]{Cornell University, School of Mechanical and Aerospace Engineering, Ithaca, United States}
+\affil[2]{Columbia University, Department of Mechanical Engineering, New York, United States}
+\affil[3]{Columbia University, Data Science Institute, New York, United States}
 \affil[*]{car45@cornell.edu}
 
-% \affil[+]{these authors contributed equally to this work}
-
-\keywords{flapping flight, unsteady aerodynamics, data mining, symbolic
-regression, evolutionary computation}
+\begin{document}
+\maketitle
 
@@
 \end{abstract}
-\begin{document}
 
-\flushbottom
-\maketitle
+\keywords{flapping flight, unsteady aerodynamics, data mining, symbolic regression, evolutionary
+computation}
+
+% \flushbottom
+% \maketitle
 % * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
@@
 %
-\thispagestyle{empty}
+% \thispagestyle{empty}
 
@@
 
-\subsection{Data-driven models}
+\subsection*{Data-driven models}
 The strategy for modeling lift forces symbolically was to train candidate equations on 58 of the 68 total experiments simultaneously, with the remaining ten reserved for testing.
@@
 
-\subsection{Analytical models for comparison}
+\subsection*{Analytical models for comparison}
 The analytical models used for comparison are also listed in Table~\ref{table:eqmodels}.
@@
 
-\subsection{Evaluation of Model Performance}
+\subsection*{Evaluation of Model Performance}
 An evaluation of model performance allows comparison between analytical models and the equations developed through the data mining process.
@@
 \includegraphics[width=\textwidth]{figures/eq_goodbad_1}
-\caption{\label{fig:eq_goodbad_1}}
+\caption{\label{fig:eq_goodbad_1} Training data.}
 \end{subfigure} &
@@
 \includegraphics[width=\textwidth]{figures/eq_goodbad_2}
-\caption{\label{fig:eq_goodbad_2}}
+\caption{\label{fig:eq_goodbad_2} Test data \#1.}
 \end{subfigure} \\
@@
 \includegraphics[width=\textwidth]{figures/eq_goodbad_3}
-\caption{\label{fig:eq_goodbad_3}}
+\caption{\label{fig:eq_goodbad_3} Test data \#2.}
 \end{subfigure} &
@@
 \includegraphics[width=\textwidth]{figures/eq_goodbad_4}
-\caption{\label{fig:eq_goodbad_4}}
+\caption{\label{fig:eq_goodbad_4} Test data \#3.}
 \end{subfigure}
@@
 
-\subsection{Performance of models trained on one experiment versus many experiments}
+\subsection*{Performance of models trained on one experiment versus many experiments}
 Figure~\ref{fig:eq_goodbad_1} illustrates the predictive performance of two models on data within
@@
 
-\subsection{Correspondence of physical terms}
+\subsection*{Correspondence of physical terms}
 One powerful result of the data-driven modeling approach is that in many cases, certain terms from the data-driven models closely match physically understandable terms from the analytical models.
@@
 \includegraphics[width=\textwidth]{figures/eq_comparison_lift}
-\caption{\label{fig:eqcomparison_lift}}
+\caption{\label{fig:eqcomparison_lift} Aerodynamic lift forces.}
 \end{subfigure}
@@
 \includegraphics[width=\textwidth]{figures/eq_comparison_inertial}
-\caption{\label{fig:eqcomparison_inertial}}
+\caption{\label{fig:eqcomparison_inertial} Inertial forces.}
 \end{subfigure}
@@
 
-\subsection{3D Printed wings and actuation mechanism}
+\subsection*{3D Printed wings and actuation mechanism}
 The experimental wings were fabricated using an Objet Connex500 3D printer with
@@
 
-\subsection{Measurement apparatus}
+\subsection*{Measurement apparatus}
 The apparatus used to measure the forces and kinematics of flapping wings was
@@
 
-\subsection{High-speed measurement of flapping kinematics}
+\subsection*{High-speed measurement of flapping kinematics}
 A Casio EXILIM EX-FH20 color video camera was used to record each experiment at
@@
 
+\bibliographystyle{abbrv}
 \bibliography{main}

### 2021-01-19 paper/main.tex
Revise introduction

@@
 with the experimental data. Then a set of data-driven equations were automatically generated from
-the data using the symbolic regression algorithm developed by Schmidt and Lipson \cite{schmidt2009distilling}.
-These data-driven models exhibit several remarkable properties.  First, they are at least as
-accurate as quasi-steady equations from the literature yet they are much simpler.  Second, they
-naturally separate distinct, understandable physical effects including terms corresponding to
-inertial forces and aerodynamic lift.  Finally, it is shown that models generated from data of a
-single experiment do not have the property of separating distinct physical effects, and they fail to
-accurately predict forces produced by a wing of a different geometry or flapping pattern.  It is
-only when models are trained on a large set of experiments including many different wings and
-flapping patterns that over-fit candidate equations are discarded and those that reflect the true
-underlying physics are maintained.  The models trained on a diverse set of data are shown to
-accurately predict the forces produced by wings of similar size and flapping kinematics that were
-not included in the training set. The ability of the symbolic regression process to infer distinct
-physical contributions to to the total lift force based on data alone suggests it may be a powerful
-tool in identifying the underlying structure of a wide variety of complex systems.
+the data using the symbolic regression algorithm developed by Schmidt and Lipson
+\cite{schmidt2009distilling}. These data-driven models exhibit several desirable properties. First,
+they are at least as accurate as quasi-steady equations from the literature yet they are much
+simpler.  Second, they naturally separate distinct, understandable physical effects including terms
+corresponding to inertial forces and aerodynamic lift. Finally, they generalize well, providing
+accurate force predictions for similar experiments outside their training dataset, indicating that
+they correctly identify and assemble the most salient and predictive variables. These results
+indicate that data-driven symbolic regression may be a powerful scientific tool in discovering the
+underlying structure of many complex systems, as well as providing a practical tool in challenging
+engineering tasks like flapping wing design optimization.
 

### 2021-01-19 paper/main.tex
Remove email
Wrap lines to 80 characters

@@
 
-\title{Automated data-driven symbolic modeling of the lift forces acting on flapping wings}
+\title{Automated data-driven symbolic modeling of the lift forces acting on
+flapping wings}
 \date{}
 
-\author[1,*]{Charles Richter}
-\author[2,3]{Hod Lipson}
-\affil[1]{Cornell University, School of Mechanical and Aerospace Engineering, Ithaca, United States}
-\affil[2]{Columbia University, Department of Mechanical Engineering, New York, United States}
-\affil[3]{Columbia University, Data Science Institute, New York, United States}
-\affil[*]{car45@cornell.edu}
+\author[1]{Charles Richter} \author[2,3]{Hod Lipson} \affil[1]{Cornell
+University, School of Mechanical and Aerospace Engineering, Ithaca, United
+States} \affil[2]{Columbia University, Department of Mechanical Engineering, New
+York, United States} \affil[3]{Columbia University, Data Science Institute, New
+York, United States}
 
@@
 
-\keywords{flapping flight, unsteady aerodynamics, data mining, symbolic regression, evolutionary
-computation}
+\keywords{flapping flight, unsteady aerodynamics, data mining, symbolic
+regression, evolutionary computation}
 
@@
 
-Quasi-steady representations of lift are particularly well suited to the process of symbolic
-regression, which harnesses computational power to conduct an automated search over the space of all
-kinematic variables and geometric parameters to find equations that most accurately predict the
-forces on a wing.  Symbolic regression algorithms are based on evolutionary computation
-\cite{koza1992genetic,forrest1993genetic} and have been used to model complex systems defined by explicit
-\cite{duffy2002using,fillon2007symbolic,bautu2005symbolic}, implicit \cite{schmidt2010symbolic}, iterated
-\cite{schmidt2009solving} and differential equations \cite{bongard2007automated}.  The
-process underlying this evolutionary computation approach is to generate an initially random
-population of equations combining time-series data and constant variables with algebraic operators,
-analytic functions and constants.  These candidate equations are then evaluated to measure their
-accuracy in predicting the measured output of the system or partial derivatives within the data.  In
-each iteration of the algorithm, the best candidates are retained and varied by recombining and
-varying their subexpressions. The algorithm is terminated when a desired level of accuracy is
-reached.  In this case, the time-series data includes the orientations, velocities and accelerations
-of wings, the constant variables include chord lengths, wingspans, centroids of wing area and other
-physical constants, and the known output of the system is the measured vertical force produced
-during flapping.  The resulting equations returned by the algorithm are quasi-steady models that can
-be compared to existing equations from the literature.
+Quasi-steady representations of lift are particularly well suited to the process
+of symbolic regression, which harnesses computational power to conduct an
+automated search over the space of all kinematic variables and geometric
+parameters to find equations that most accurately predict the forces on a wing.
+Symbolic regression algorithms are based on evolutionary computation
+\cite{koza1992genetic,forrest1993genetic} and have been used to model complex
+systems defined by explicit
+\cite{duffy2002using,fillon2007symbolic,bautu2005symbolic}, implicit
+\cite{schmidt2010symbolic}, iterated \cite{schmidt2009solving} and differential
+equations \cite{bongard2007automated}.  The process underlying this evolutionary
+computation approach is to generate an initially random population of equations
+combining time-series data and constant variables with algebraic operators,
+analytic functions and constants.  These candidate equations are then evaluated
+to measure their accuracy in predicting the measured output of the system or
+partial derivatives within the data.  In each iteration of the algorithm, the
+best candidates are retained and varied by recombining and varying their
+subexpressions. The algorithm is terminated when a desired level of accuracy is
+reached.  In this case, the time-series data includes the orientations,
+velocities and accelerations of wings, the constant variables include chord
+lengths, wingspans, centroids of wing area and other physical constants, and the
+known output of the system is the measured vertical force produced during
+flapping.  The resulting equations returned by the algorithm are quasi-steady
+models that can be compared to existing equations from the literature.
 
@@
 
-Numerous studies have measured the forces produced by a flapping wing in
-various experimental setups including flapping in a liquid bath such as oil
+Numerous studies have measured the forces produced by a flapping wing in various
+experimental setups including flapping in a liquid bath such as oil
 \cite{walker2009photogrammetric}, by tracking the kinematics of insects in
 flight \cite{fry2003aerodynamics,ristroph2009automated} and by tracking the
-detailed wing motions and visualizing the airflow around the wings of insects
-in a wind tunnel \cite{young2009details}.  Only recently have studies directly
+detailed wing motions and visualizing the airflow around the wings of insects in
+a wind tunnel \cite{young2009details}.  Only recently have studies directly
 measured the forces involved in flapping with at-scale robotic wings or actual
@@
 
-Here we present a series of experiments measuring the kinematics and lift forces produced by eleven
-differently sized wings flapping at multiple speeds and amplitudes with Reynolds numbers between
-2500 and 5000. Several quasi-steady models from the literature are shown to exhibit strong agreement
-with the experimental data. Then a set of data-driven equations were automatically generated from
-the data using the symbolic regression algorithm developed by Schmidt and Lipson
-\cite{schmidt2009distilling}. These data-driven models exhibit several desirable properties. First,
-they are at least as accurate as quasi-steady equations from the literature yet they are much
-simpler.  Second, they naturally separate distinct, understandable physical effects including terms
-corresponding to inertial forces and aerodynamic lift. Finally, they generalize well, providing
-accurate force predictions for similar experiments outside their training dataset, indicating that
-they correctly identify and assemble the most salient and predictive variables. These results
-indicate that data-driven symbolic regression may be a powerful scientific tool in discovering the
-underlying structure of many complex systems, as well as providing a practical tool in challenging
-engineering tasks like flapping wing design optimization.
+Here we present a series of experiments measuring the kinematics and lift forces
+produced by eleven differently sized wings flapping at multiple speeds and
+amplitudes with Reynolds numbers between 2500 and 5000. Several quasi-steady
+models from the literature are shown to exhibit strong agreement with the
+experimental data. Then a set of data-driven equations were automatically
+generated from the data using the symbolic regression algorithm developed by
+Schmidt and Lipson \cite{schmidt2009distilling}. These data-driven models
+exhibit several desirable properties. First, they are at least as accurate as
+quasi-steady equations from the literature yet they are much simpler.  Second,
+they naturally separate distinct, understandable physical effects including
+terms corresponding to inertial forces and aerodynamic lift. Finally, they
+generalize well, providing accurate force predictions for similar experiments
+outside their training dataset, indicating that they correctly identify and
+assemble the most salient and predictive variables. These results indicate that
+data-driven symbolic regression may be a powerful scientific tool in discovering
+the underlying structure of many complex systems, as well as providing a
+practical tool in challenging engineering tasks like flapping wing design
+optimization.
 
@@
 \subsection*{Data-driven models}
-The strategy for modeling lift forces symbolically was to train candidate equations on 58 of the 68 total experiments simultaneously, with the remaining ten reserved for testing.
-These 58 training sets were intended to represent a diverse sampling of wing geometry, flapping speed and amplitude combinations.
-The results from these 58 experiments were compiled into a single dataset listing roughly two periods of smoothed, registered flapping data per experiment.
-The data consisted of instantaneous measured vertical force along with the measured value of each kinematic variable, the first and second derivatives of those variables, and the geometric parameters necessary to describe the wing's shape and distribution of area, totaling 31 separate variables.
+The strategy for modeling lift forces symbolically was to train candidate
+equations on 58 of the 68 total experiments simultaneously, with the remaining
+ten reserved for testing.  These 58 training sets were intended to represent a
+diverse sampling of wing geometry, flapping speed and amplitude combinations.
+The results from these 58 experiments were compiled into a single dataset
+listing roughly two periods of smoothed, registered flapping data per
+experiment.  The data consisted of instantaneous measured vertical force along
+with the measured value of each kinematic variable, the first and second
+derivatives of those variables, and the geometric parameters necessary to
+describe the wing's shape and distribution of area, totaling 31 separate
+variables.
 
-By training on 58 experiments simultaneously, the symbolic regression algorithm generated equations that fit all of the data, using the some, but not all, of the provided variables to account for variations in kinematics and wing geometry between experiments.
-Table~\ref{table:eqmodels} lists five of the leading candidate models, which were trained using mean absolute error (MAE) as the fitness metric during the search.
-The search was conducted for approximately 24 hours, using 60 cores on a computing cluster.
+By training on 58 experiments simultaneously, the symbolic regression algorithm
+generated equations that fit all of the data, using the some, but not all, of
+the provided variables to account for variations in kinematics and wing geometry
+between experiments.   Table~\ref{table:eqmodels} lists five of the leading
+candidate models, which were trained using mean absolute error (MAE) as the
+fitness metric during the search. The search was conducted for approximately 24
+hours, using 60 cores on a computing cluster.
 
@@
 \end{tabular}
-\caption{\label{table:eqmodels}Comparison of data-driven models and analytical equations of lift.
-The models are sorted by equation size, which is calculated as the sum of arithmetic and
-trigonometric operators required to compute a force prediction from kinematic and geometric data.
-MAE reports average mean absolute error (in Newtons of vertical force) for the ten test
+\caption{\label{table:eqmodels}Comparison of data-driven models and analytical
+  equations of lift.  The models are sorted by equation size, which is
+  calculated as the sum of arithmetic and trigonometric operators required to
+  compute a force prediction from kinematic and geometric data.  MAE reports
+  average mean absolute error (in Newtons of vertical force) for the ten test
   experiments.}
@@
 \subsection*{Analytical models for comparison}
-The analytical models used for comparison are also listed in Table~\ref{table:eqmodels}.
-These models consist of calculated inertial forces added to a collection of aerodynamic terms.
-The primary aerodynamic contribution in all quasi-steady models is a translational lift term proportional to the squared velocity of each span-element of the wing that is related to the wing's angle of attack and empirically determined coefficient of lift.
-This term is labeled $L_{planar}$ when wing motions out of the nominal horizontal flapping plane are neglected in the calculations, or $L_{trans}$ when out-of-plane motions are included.
-This translational lift term is given by:
+The analytical models used for comparison are also listed in
+Table~\ref{table:eqmodels}.  These models consist of calculated inertial forces
+added to a collection of aerodynamic terms.  The primary aerodynamic
+contribution in all quasi-steady models is a translational lift term
+proportional to the squared velocity of each span-element of the wing that is
+related to the wing's angle of attack and empirically determined coefficient of
+lift.  This term is labeled $L_{planar}$ when wing motions out of the nominal
+horizontal flapping plane are neglected in the calculations, or $L_{trans}$ when
+out-of-plane motions are included. This translational lift term is given by:
 \begin{equation}
@@
 
-Other aerodynamic contributors include rotational lift proportional to the product of a wing's rotational and translational velocities, labeled $L_{rot}$, as well as "added-mass" effects related to the accelerations of fluid surrounding the wing \cite{sedov1965two}.
-These added mass terms are proportional to the wing's linear acceleration ($AM_{trans}$), rotational acceleration ($AM_{rot}$), or a product of the two ($AM_{trans-rot}$).
-Two sets of quasi-steady equations from the literature include various combinations of such terms \cite{whitney2010aeromechanics,pesavento2004falling}, which allow more flexible model fitting through additional coefficients.
+Other aerodynamic contributors include rotational lift proportional to the
+product of a wing's rotational and translational velocities, labeled $L_{rot}$,
+as well as "added-mass" effects related to the accelerations of fluid
+surrounding the wing \cite{sedov1965two}.  These added mass terms are
+proportional to the wing's linear acceleration ($AM_{trans}$), rotational
+acceleration ($AM_{rot}$), or a product of the two ($AM_{trans-rot}$).  Two sets
+of quasi-steady equations from the literature include various combinations of
+such terms \cite{whitney2010aeromechanics,pesavento2004falling}, which allow
+more flexible model fitting through additional coefficients.
 
 Inertial forces acting on the wings are modeled as a sum of two contributions.
-The primary contribution results from the pendulum-like swinging motion that occurs when the deflection angle ($\psi$) flips at stroke reversal.
-The secondary contribution results from small angle deviations from the horizontal flapping plane ($\theta$).
-These angles are depicted in Figure~\ref{fig:apparatus}.
-These inertial forces are given by:
+The primary contribution results from the pendulum-like swinging motion that
+occurs when the deflection angle ($\psi$) flips at stroke reversal.  The
+secondary contribution results from small angle deviations from the horizontal
+flapping plane ($\theta$).  These angles are depicted in
+Figure~\ref{fig:apparatus}.  These inertial forces are given by:
 \begin{equation}\label{sdbla}
@@
 \end{subfigure}
-\caption{\subref{fig:apparatus} Schematic diagram of the experimental apparatus showing three
-  kinematic variables of the wing motion: stroke ($\phi$), deflection ($\psi$) and deviation
-  ($\theta$). \subref{fig:pareto_mae} Pareto plot of all data-driven (``EQ Models'') and analytical
-  models using mean values for model coefficients. All data-driven models were smaller and more
+\caption{\subref{fig:apparatus} Schematic diagram of the experimental apparatus
+  showing three kinematic variables of the wing motion: stroke ($\phi$),
+  deflection ($\psi$) and deviation ($\theta$). \subref{fig:pareto_mae} Pareto
+  plot of all data-driven (``EQ Models'') and analytical models using mean
+  values for model coefficients. All data-driven models were smaller and more
   accurate than analytical models.
@@
 \subsection*{Evaluation of Model Performance}
-An evaluation of model performance allows comparison between analytical models and the equations developed through the data mining process.
-Equations are judged by their accuracy and complexity.
-In this case, the complexity or size of an equation was measured by simply counting the number of arithmetic and trigonometric operators used to combine the measured kinematic variables and geometric properties into an equation for instantaneous vertical force.
-Both our data-driven models and the analytical equations are measured in this way.
-For fairness, constants in the analytical models that would have been lumped together into larger coefficients by the symbolic regression algorithm (such as material density, air density, \(\pi\) or factors of integration) are also lumped together in the calculations of size for the analytical equations.
-The size of the inertial calculations was added to the sizes of analytical lift calculations.
-The
-pendulum motion has an equation size of 13 and the total inertial calculation including out-of-plane deviation has a size of 27.
+An evaluation of model performance allows comparison between analytical models
+and the equations developed through the data mining process.  Equations are
+judged by their accuracy and complexity.  In this case, the complexity or size
+of an equation was measured by simply counting the number of arithmetic and
+trigonometric operators used to combine the measured kinematic variables and
+geometric properties into an equation for instantaneous vertical force.  Both
+our data-driven models and the analytical equations are measured in this way.
+For fairness, constants in the analytical models that would have been lumped
+together into larger coefficients by the symbolic regression algorithm (such as
+material density, air density, \(\pi\) or factors of integration) are also
+lumped together in the calculations of size for the analytical equations. The
+size of the inertial calculations was added to the sizes of analytical lift
+calculations.  The pendulum motion has an equation size of 13 and the total
+inertial calculation including out-of-plane deviation has a size of 27.
 
-Table~\ref{table:eqmodels} shows the size of all equations considered in this study, and indicates that the analytical models are all significantly more complex than our data-driven models.
-One of the reasons for high complexity in the analytical models is the dependence on wing geometry and integration of forces over the span and chord of the wing.
-A second reason for the complexity in analytical models is the number of geometric and coordinate transformations required to compute the necessary linear and rotational velocities.
-In this sense, several terms generated during symbolic regression can be considered to be approximations of much more subtle calculations of kinematic quantities.
+Table~\ref{table:eqmodels} shows the size of all equations considered in this
+study, and indicates that the analytical models are all significantly more
+complex than our data-driven models.  One of the reasons for high complexity in
+the analytical models is the dependence on wing geometry and integration of
+forces over the span and chord of the wing.  A second reason for the complexity
+in analytical models is the number of geometric and coordinate transformations
+required to compute the necessary linear and rotational velocities.  In this
+sense, several terms generated during symbolic regression can be considered to
+be approximations of much more subtle calculations of kinematic quantities.
 
-The accuracy of each equation was measured as the mean absolute error (MAE) between the predicted and actual vertical force (in Newtons) produced by the flapping wing.
-Figure~\ref{fig:pareto_mae} shows the Pareto plot of all models considered, depicting accuracy as a function of model complexity.
-It is important to note that all data-driven models have between three and five coefficients, so tuning these coefficients allows further minimization of error compared to the analytical models, which have between one and three.
-Nevertheless, discovery of models with smaller error than theoretical predictions that are simultaneously much simpler is a significant feat for the data mining process and symbolic regression.
+The accuracy of each equation was measured as the mean absolute error (MAE)
+between the predicted and actual vertical force (in Newtons) produced by the
+flapping wing.  Figure~\ref{fig:pareto_mae} shows the Pareto plot of all models
+considered, depicting accuracy as a function of model complexity. It is
+important to note that all data-driven models have between three and five
+coefficients, so tuning these coefficients allows further minimization of error
+compared to the analytical models, which have between one and three.
+Nevertheless, discovery of models with smaller error than theoretical
+predictions that are simultaneously much simpler is a significant feat for the
+data mining process and symbolic regression.
 
@@
 \subsection*{Performance of models trained on one experiment versus many experiments}
-Figure~\ref{fig:eq_goodbad_1} illustrates the predictive performance of two models on data within
-their training sets, while Figures~\ref{fig:eq_goodbad_2},~\ref{fig:eq_goodbad_3},
-and~\ref{fig:eq_goodbad_4} illustrate the performance of those same two models on data outside
-their training sets. The differences between these two models result from the fact that one of them,
-$EQ_4$ from Table~\ref{table:eqmodels} (drawn in blue), was trained on 58 separate experiments
-simultaneously, whereas the other one, shown below (drawn in red), is one of many models trained
-only on the single experiment shown in Figure~\ref{fig:eq_goodbad_1}:
+Figure~\ref{fig:eq_goodbad_1} illustrates the predictive performance of two
+models on data within their training sets, while
+Figures~\ref{fig:eq_goodbad_2},~\ref{fig:eq_goodbad_3},
+and~\ref{fig:eq_goodbad_4} illustrate the performance of those same two models
+on data outside their training sets. The differences between these two models
+result from the fact that one of them, $EQ_4$ from Table~\ref{table:eqmodels}
+(drawn in blue), was trained on 58 separate experiments simultaneously, whereas
+the other one, shown below (drawn in red), is one of many models trained only on
+the single experiment shown in Figure~\ref{fig:eq_goodbad_1}:
 \begin{align}
@@
 \end{align}
-Thus, $EQ_4$ shows strong qualitative and quantitative agreement with all of the experiments in Figure~\ref{fig:eq_goodbad}, indicating that this model is not over-fit but in fact generalizes well to new data.
-Equation~\eqref{eq:badeq}, on the other hand, fails to accurately predict the recorded data outside its training set.
-Insufficient data and over-fitting are potential pitfalls of data-driven modeling, but are overcome in this case with a diverse sampling of training data.
+Thus, $EQ_4$ shows strong qualitative and quantitative agreement with all of the
+experiments in Figure~\ref{fig:eq_goodbad}, indicating that this model is not
+over-fit but in fact generalizes well to new data. Equation~\eqref{eq:badeq}, on
+the other hand, fails to accurately predict the recorded data outside its
+training set.  Insufficient data and over-fitting are potential pitfalls of
+data-driven modeling, but are overcome in this case with a diverse sampling of
+training data.
 
-Since Equation~\eqref{eq:badeq} is unable to accurately predict forces outside its training set, it is unlikely to offer any insight into the underlying physical processes.
-$EQ_4$, however, must capture some of the underlying structure of the physics in order to generalize to data upon which it was not trained.
-In data-driven modeling, training on many experiments simultaneously helps to reject candidate equations that may fit some, but not all, of the data, and helps to retain those models that reflect the underlying physics and offer insight into the structure of the additive components that compose the observed behavior of a complex dynamical system.
+Since Equation~\eqref{eq:badeq} is unable to accurately predict forces outside
+its training set, it is unlikely to offer any insight into the underlying
+physical processes.  $EQ_4$, however, must capture some of the underlying
+structure of the physics in order to generalize to data upon which it was not
+trained.  In data-driven modeling, training on many experiments simultaneously
+helps to reject candidate equations that may fit some, but not all, of the data,
+and helps to retain those models that reflect the underlying physics and offer
+insight into the structure of the additive components that compose the observed
+behavior of a complex dynamical system.
 
 \subsection*{Correspondence of physical terms}
-One powerful result of the data-driven modeling approach is that in many cases, certain terms from the data-driven models closely match physically understandable terms from the analytical models.
-The correspondence between these terms is listed in Table~\ref{table:eqcorrespondence}.
-Taking model $EQ_4$ as an example, all three terms show strong resemblance to inertial and aerodynamic components of the analytical models.
-Figure~\ref{fig:eqcomparison} shows a comparison between the simple velocity-squared aerodynamic lift prediction and the second term of $EQ_4$: \(C_2RX_{CM}\psi\dot{\phi}\).
-These two calculations are numerically very similar, yet the third term of $EQ_4$ is much simpler than $L_{trans}$.
+One powerful result of the data-driven modeling approach is that in many cases,
+certain terms from the data-driven models closely match physically
+understandable terms from the analytical models.  The correspondence between
+these terms is listed in Table~\ref{table:eqcorrespondence}. Taking model $EQ_4$
+as an example, all three terms show strong resemblance to inertial and
+aerodynamic components of the analytical models.  Figure~\ref{fig:eqcomparison}
+shows a comparison between the simple velocity-squared aerodynamic lift
+prediction and the second term of $EQ_4$: \(C_2RX_{CM}\psi\dot{\phi}\).  These
+two calculations are numerically very similar, yet the third term of $EQ_4$ is
+much simpler than $L_{trans}$.
 
-Similarly, the sum of terms 1 and 3, \(C_1RY_{CM}\dot{\psi}^2\cos(\psi) - C_3R\ddot{\theta}\), closely approximates the calculation of inertial forces including the pendulum motion of the wing as well as the out-of-plane deviation.
-Figure~\ref{fig:eqcomparison} also shows these two calculations, indicating very similar structure.
-While $EQ_4$'s approximation of these inertial forces does not precisely match the analytical calculation, it follows a very similar pattern and is much simpler.
-Remarkably, these inertial terms utilize some of the same variable combinations as the analytical calculations to express the inertial force, namely \(Y_{CM}\dot{\psi}^2\cos(\psi)\) and \(\ddot{\theta}\), with different coefficients to express the geometric variation between wings and small angle approximations \(\cos\theta\approx1\), \(\sin\theta\approx0\) for small out-of-plane deviation angles.
-Since \(|\theta|<2^{\circ}\) in general, the small angle approximation is valid.
+Similarly, the sum of terms 1 and 3, \(C_1RY_{CM}\dot{\psi}^2\cos(\psi) -
+C_3R\ddot{\theta}\), closely approximates the calculation of inertial forces
+including the pendulum motion of the wing as well as the out-of-plane deviation.
+Figure~\ref{fig:eqcomparison} also shows these two calculations, indicating very
+similar structure.  While $EQ_4$'s approximation of these inertial forces does
+not precisely match the analytical calculation, it follows a very similar
+pattern and is much simpler.  Remarkably, these inertial terms utilize some of
+the same variable combinations as the analytical calculations to express the
+inertial force, namely \(Y_{CM}\dot{\psi}^2\cos(\psi)\) and \(\ddot{\theta}\),
+with different coefficients to express the geometric variation between wings and
+small angle approximations \(\cos\theta\approx1\), \(\sin\theta\approx0\) for
+small out-of-plane deviation angles.  Since \(|\theta|<2^{\circ}\) in general,
+the small angle approximation is valid.
 
@@
 \end{subfigure}
-\caption{Comparison between terms of the analytical expressions (red) and data-driven models (blue).
-  \subref{fig:eqcomparison_lift} Primary translational aerodynamic lift proportional to the square
-  of the wing velocity. \subref{fig:eqcomparison_inertial} Inertial forces related to the pendulum
-  motion of the wing and deviation out of the stroke plane. The modeling process automatically
-  decoupled aerodynamic and inertial effects with no prior knowledge of the system. The total
-  measured force is shown for reference (black).}
+\caption{Comparison between terms of the analytical expressions (red) and
+  data-driven models (blue).  \subref{fig:eqcomparison_lift} Primary
+  translational aerodynamic lift proportional to the square of the wing
+  velocity. \subref{fig:eqcomparison_inertial} Inertial forces related to the
+  pendulum motion of the wing and deviation out of the stroke plane. The
+  modeling process automatically decoupled aerodynamic and inertial effects with
+  no prior knowledge of the system. The total measured force is shown for
+  reference (black).}
 \label{fig:eqcomparison}
@@
 Table~\ref{table:eqcorrespondence} lists the corresponding terms between the
-basic analytical expressions and the data-driven model $EQ_4$.  While two of
-the inertial terms share the same combinations of variables, the data-driven
+basic analytical expressions and the data-driven model $EQ_4$.  While two of the
+inertial terms share the same combinations of variables, the data-driven
 expression for aerodynamic lift is very different from its analytical
@@
 level of accuracy.  Therefore, key physical effects may be represented
-numerically by terms that are symbolically different and often much simpler
-than their analytical counterparts.  This behavior of automatically separating
+numerically by terms that are symbolically different and often much simpler than
+their analytical counterparts.  This behavior of automatically separating
 distinct dynamical building blocks is one of the most powerful aspects of
@@
 was numerically similar to the complete calculation of the pendulum and
-out-of-plane motions with some analytical terms in common.  Finally, by
-training data-driven models on experimental data from 58 individual
-wing-amplitude-speed combinations simultaneously, the symbolic-regression
-algorithm was able to discard models that did not generalize, while retaining
-those whose terms captured the underlying physical processes.
+out-of-plane motions with some analytical terms in common.  Finally, by training
+data-driven models on experimental data from 58 individual wing-amplitude-speed
+combinations simultaneously, the symbolic-regression algorithm was able to
+discard models that did not generalize, while retaining those whose terms
+captured the underlying physical processes.
 
@@
 could be useful in phenomenological modeling and guiding the exploration of a
-wide variety of complex systems.  Its ability to automatically identify
-distinct physical effects that are consistent across all datasets makes it a
-powerful tool for identifying the structure of contributions from building
-blocks that make up the complete observed behavior of a dynamical system.
+wide variety of complex systems.  Its ability to automatically identify distinct
+physical effects that are consistent across all datasets makes it a powerful
+tool for identifying the structure of contributions from building blocks that
+make up the complete observed behavior of a dynamical system.
 
@@
 The experimental wings were fabricated using an Objet Connex500 3D printer with
-the rigid Objet FullCure720 material.
-Flapping flight with 3D printed wings was reported in \cite{richter2011untethered}. The
-wing design consisted of an interlocking root and tip, allowing the wing
-surface to pivot naturally during flapping from its neutral vertical position
-to an angle of attack of approximately 45\textdegree. The wing design is illustrated in
-Figure~\ref{fig:wing1} and the hinge design is
-illustrated in Figure~\ref{fig:root2}.  In total, eleven wings were used,
-ranging in wingspan from 40 mm to 120 mm and ranging in chord length from 40 mm
-to 80 mm. Flapping amplitudes were constrained by the geometry of the wing root
-design to be set at 72\textdegree, 81\textdegree, and 96\textdegree, and each
-wing was flapped at speeds ranging between 1 Hz and 3.5 Hz for each of these
-amplitude settings.  68 sets of data were recorded, representing Reynolds
-numbers between 2500 and 5000.
+the rigid Objet FullCure720 material.  Flapping flight with 3D printed wings was
+reported in \cite{richter2011untethered}. The wing design consisted of an
+interlocking root and tip, allowing the wing surface to pivot naturally during
+flapping from its neutral vertical position to an angle of attack of
+approximately 45\textdegree. The wing design is illustrated in
+Figure~\ref{fig:wing1} and the hinge design is illustrated in
+Figure~\ref{fig:root2}.  In total, eleven wings were used, ranging in wingspan
+from 40 mm to 120 mm and ranging in chord length from 40 mm to 80 mm. Flapping
+amplitudes were constrained by the geometry of the wing root design to be set at
+72\textdegree, 81\textdegree, and 96\textdegree, and each wing was flapped at
+speeds ranging between 1 Hz and 3.5 Hz for each of these amplitude settings.  68
+sets of data were recorded, representing Reynolds numbers between 2500 and 5000.
 
@@
 sinusoidal flapping motion. This system is pictured in
-Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency
-of the load cell and wing driving mechanism was roughly 70 Hz, the natural
+Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency of
+the load cell and wing driving mechanism was roughly 70 Hz, the natural
 frequency for the first bending mode of the experimental wing was approximately
@@
 digital low-pass filter with a cutoff frequency of 7.5 times the flapping
-frequency, which preserved the primary inertial and aerodynamic forces acting
-on the wings while eliminating the faster vibrations of the wing structure and
-the load cell apparatus.
+frequency, which preserved the primary inertial and aerodynamic forces acting on
+the wings while eliminating the faster vibrations of the wing structure and the
+load cell apparatus.
 
@@
 \end{subfigure}
-\caption{ \subref{fig:apparatus2} Front view of the flapping wing apparatus showing the wing mounted
-  on its driving mechanism attached to the force sensor. \subref{fig:apparatus3} Close-up view of
-  the 3D printed frame and crank mechanism used to drive the flapping motion of the wing.}
+\caption{\subref{fig:apparatus2} Front view of the flapping wing apparatus
+  showing the wing mounted on its driving mechanism attached to the force
+  sensor. \subref{fig:apparatus3} Close-up view of the 3D printed frame and
+  crank mechanism used to drive the flapping motion of the wing.}
 \end{figure}
@@
 was colored red to enable precise tracking of the stroke angle.  The top-down
-projected area of the wing was used to extract deflection angle, and tracking
-of the wing tip in the side-view was used to extract the angle of deviation out
-of the wing's stroke plane.  These three angles precisely define the
-configuration of the wing.  A schematic diagram of this apparatus is shown in
+projected area of the wing was used to extract deflection angle, and tracking of
+the wing tip in the side-view was used to extract the angle of deviation out of
+the wing's stroke plane.  These three angles precisely define the configuration
+of the wing.  A schematic diagram of this apparatus is shown in
 Figure~\ref{fig:apparatus}. A variety of image-processing techniques were used
@@
 red-colored main beam was identified with a color filter and a line was fitted
-using a Hough Transform to identify the stroke angle.  The wing deflection
-angle was computed by measuring the projected area of the wing in the image and
+using a Hough Transform to identify the stroke angle.  The wing deflection angle
+was computed by measuring the projected area of the wing in the image and
 compared with the known dimensions of the rectangular wing shape. The
 out-of-plane deviation angle of the wing was measured by tracking the leading
-and trailing edge corners of the wingtip in the side-view of the mirror. We refer the reader to our
-previous work for a more detailed description of our methods of extracting wing kinematics from
-high speed video and processing the measured data\cite{richter2012thesis}.
-% Figure~\ref{fig:kinematics} provides sample force data and measured wing kinematics from one
-% experiment.
+and trailing edge corners of the wingtip in the side-view of the mirror. We
+refer the reader to our previous work for a more detailed description of our
+methods of extracting wing kinematics from high speed video and processing the
+measured data\cite{richter2012thesis}.
+% Figure~\ref{fig:kinematics} provides sample force data and measured wing
+% kinematics from one experiment.
 

### 2021-01-19 paper/main.tex
Move Methods section after Introduction
Move schematic diagram into Methods section and add introductory paragraph to Methods section

@@
 
+\section*{Methods}
+Our experimental apparatus consisted of a wing and actuator mounted to a
+vertical force sensor, with a high-speed video camera recording the motion of
+the wing from a top-down view and (using a mirror) also from a side-view.
+Figure~\ref{fig:apparatus} provides a diagram of the apparatus as well as
+illustrations of the three primary variables describing the orientation of the
+wing in the apparatus: Stroke ($\phi$), deflection ($\psi$) and deviation
+($\theta$). The stroke motion is enabled by a hinge joint and is the only
+actuated degree of freedom in the system. The deflection of the wing is enabled
+by a hinge in the root of the wing, and is a passive motion resulting from both
+inertial and aerodynamic effects. Deviation refers to the passive vertical
+motion of the wing out of its horizontal flapping plane, and is enabled only by
+the tolerances of the hinges and the flexibility of the wing structure itself.
+
+% Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.
+
+\begin{figure}
+\centering
+\includegraphics[trim=0 15 0 5, clip, width=0.5\textwidth]{figures/apparatus}
+\caption{Schematic diagram of the experimental apparatus
+  showing three kinematic variables of the wing motion: stroke ($\phi$),
+  deflection ($\psi$) and deviation ($\theta$).\label{fig:apparatus}}
+\end{figure}
+
+\subsection*{3D Printed wings and actuation mechanism}
+The experimental wings were fabricated using an Objet Connex500 3D printer with
+the rigid Objet FullCure720 material.  Flapping flight with 3D printed wings was
+reported in \cite{richter2011untethered}. The wing design consisted of an
+interlocking root and tip, allowing the wing surface to pivot naturally during
+flapping from its neutral vertical position to an angle of attack of
+approximately 45\textdegree. The wing design is illustrated in
+Figure~\ref{fig:wing1} and the hinge design is illustrated in
+Figure~\ref{fig:root2}.  In total, eleven wings were used, ranging in wingspan
+from 40 mm to 120 mm and ranging in chord length from 40 mm to 80 mm. Flapping
+amplitudes were constrained by the geometry of the wing root design to be set at
+72\textdegree, 81\textdegree, and 96\textdegree, and each wing was flapped at
+speeds ranging between 1 Hz and 3.5 Hz for each of these amplitude settings.  68
+sets of data were recorded, representing Reynolds numbers between 2500 and 5000.
+
+% See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
+\begin{figure}[h]%[!htb]
+\centering
+\begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[width=0.95\textwidth]{figures/wing1}
+    \caption{\label{fig:wing1}}
+\end{subfigure}
+\begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[trim=0 105 0 0, clip, width=0.95\textwidth]{figures/root2}
+    \caption{\label{fig:root2}}
+\end{subfigure}
+\caption{Detail of 3D printed wing \subref{fig:wing1} and wing root \subref{fig:root2} design
+  showing concentric-cylindrical design to allow passive pivoting of the wing surface.}
+\end{figure}
+
+\subsection*{Measurement apparatus}
+The apparatus used to measure the forces and kinematics of flapping wings was
+based on a FUTEK LSM250 parallelogram load cell with a USB interface for data
+logging.  This force measurement system was capable of a sampling rate of 1 kHz
+and a maximum measurement precision of 1\e{-9} N.  Mounted directly to the load
+cell was a wing-flapping mechanism, which consisted of a DC gear motor mated to
+a crank and connecting rod used to drive the experimental wings in a roughly
+sinusoidal flapping motion. This system is pictured in
+Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency of
+the load cell and wing driving mechanism was roughly 70 Hz, the natural
+frequency for the first bending mode of the experimental wing was approximately
+26 Hz, and experimental wings were flapped at or below approximately 3.5 Hz.
+All experimental force measurements were filtered using a high-order zero-phase
+digital low-pass filter with a cutoff frequency of 7.5 times the flapping
+frequency, which preserved the primary inertial and aerodynamic forces acting on
+the wings while eliminating the faster vibrations of the wing structure and the
+load cell apparatus.
+
+% % See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
+\begin{figure}%[!htb]
+\centering
+\begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[height=5cm]{figures/apparatus2}
+    \caption{\label{fig:apparatus2}}
+\end{subfigure}
+\begin{subfigure}[t]{0.45\textwidth}
+    \centering
+    \includegraphics[height=5cm]{figures/apparatus3}
+    \caption{\label{fig:apparatus3}}
+\end{subfigure}
+\caption{\subref{fig:apparatus2} Front view of the flapping wing apparatus
+  showing the wing mounted on its driving mechanism attached to the force
+  sensor. \subref{fig:apparatus3} Close-up view of the 3D printed frame and
+  crank mechanism used to drive the flapping motion of the wing.}
+\end{figure}
+
+\subsection*{High-speed measurement of flapping kinematics}
+A Casio EXILIM EX-FH20 color video camera was used to record each experiment at
+210 frames per second from a top-down perspective as well as a side view
+provided by a mirror mounted in the field of view.  The main beam of each wing
+was colored red to enable precise tracking of the stroke angle.  The top-down
+projected area of the wing was used to extract deflection angle, and tracking of
+the wing tip in the side-view was used to extract the angle of deviation out of
+the wing's stroke plane.  These three angles precisely define the configuration
+of the wing.  A schematic diagram of this apparatus is shown in
+Figure~\ref{fig:apparatus}. A variety of image-processing techniques were used
+to extract the kinematics of the wing motion from high-speed video.  The
+red-colored main beam was identified with a color filter and a line was fitted
+using a Hough Transform to identify the stroke angle.  The wing deflection angle
+was computed by measuring the projected area of the wing in the image and
+compared with the known dimensions of the rectangular wing shape. The
+out-of-plane deviation angle of the wing was measured by tracking the leading
+and trailing edge corners of the wingtip in the side-view of the mirror. We
+refer the reader to our previous work for a more detailed description of our
+methods of extracting wing kinematics from high speed video and processing the
+measured data\cite{richter2012thesis}.
+% Figure~\ref{fig:kinematics} provides sample force data and measured wing
+% kinematics from one experiment.
+
+% \begin{figure}[ht]
+% \centering
+% \includegraphics[width=\linewidth]{figures/stream}
+% \caption{Legend (350 words max). Example legend text.}
+% \label{fig:stream}
+% \end{figure}
+
+% \begin{figure}
+% \centering
+% \begin{subfigure}[b]{.45\linewidth}
+% \includegraphics[width=\linewidth]{figures/regforce1}
+% \caption{Measured vertical force.}
+% \label{fig:regforce}
+% \end{subfigure}
+% \begin{subfigure}[b]{.45\linewidth}
+% \includegraphics[width=\linewidth]{figures/regstroke1}
+% \caption{Wing stroke angle.}
+% \label{fig:regstroke}
+% \end{subfigure}
+
+% \begin{subfigure}[b]{.45\linewidth}
+% \includegraphics[width=\linewidth]{figures/regdef1}
+% \caption{Wing deflection angle.}
+% \label{fig:regdef}
+% \end{subfigure}
+% \begin{subfigure}[b]{.45\linewidth}
+% \includegraphics[width=\linewidth]{figures/regdev1}
+% \caption{Wing deviation angle.}
+% \label{fig:regdev}
+% \end{subfigure}
+% \caption{Measured \subref{fig:regforce} vertical force, \subref{fig:regstroke} wing stroke angle,
+%   \subref{fig:regdef} deflection angle, \subref{fig:regdev} and deviation angle from one experiment,
+%   registered and averaged over 18 flapping periods.}
+% \label{fig:kinematics}
+% \end{figure}
+
 \section*{Results}
@@
 
-\begin{figure}%[!htb]
+\begin{figure}
 \centering
-\begin{subfigure}{0.3\textwidth}
-\includegraphics[trim=0 15 0 5, clip, width=\textwidth]{figures/apparatus}
-\caption{\label{fig:apparatus}}
-\end{subfigure}
-\qquad
-\begin{subfigure}{0.5\textwidth}
-\centering
-\includegraphics[width=\textwidth]{figures/mae_nofit}
-\caption{\label{fig:pareto_mae}}
-\end{subfigure}
-\caption{\subref{fig:apparatus} Schematic diagram of the experimental apparatus
-  showing three kinematic variables of the wing motion: stroke ($\phi$),
-  deflection ($\psi$) and deviation ($\theta$). \subref{fig:pareto_mae} Pareto
-  plot of all data-driven (``EQ Models'') and analytical models using mean
-  values for model coefficients. All data-driven models were smaller and more
-  accurate than analytical models.
-}
+\includegraphics[width=0.5\textwidth]{figures/mae_nofit}
+\caption{Pareto plot of all data-driven (``EQ Models'') and analytical models
+  using mean values for model coefficients. All data-driven models were smaller
+  and more accurate than analytical models.\label{fig:pareto_mae}}
 \end{figure}
@@
 
-\section*{Methods}
-
-% Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.
-
-\subsection*{3D Printed wings and actuation mechanism}
-The experimental wings were fabricated using an Objet Connex500 3D printer with
-the rigid Objet FullCure720 material.  Flapping flight with 3D printed wings was
-reported in \cite{richter2011untethered}. The wing design consisted of an
-interlocking root and tip, allowing the wing surface to pivot naturally during
-flapping from its neutral vertical position to an angle of attack of
-approximately 45\textdegree. The wing design is illustrated in
-Figure~\ref{fig:wing1} and the hinge design is illustrated in
-Figure~\ref{fig:root2}.  In total, eleven wings were used, ranging in wingspan
-from 40 mm to 120 mm and ranging in chord length from 40 mm to 80 mm. Flapping
-amplitudes were constrained by the geometry of the wing root design to be set at
-72\textdegree, 81\textdegree, and 96\textdegree, and each wing was flapped at
-speeds ranging between 1 Hz and 3.5 Hz for each of these amplitude settings.  68
-sets of data were recorded, representing Reynolds numbers between 2500 and 5000.
-
-% See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
-\begin{figure}[h]%[!htb]
-\centering
-\begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \includegraphics[width=0.95\textwidth]{figures/wing1}
-    \caption{\label{fig:wing1}}
-\end{subfigure}
-\begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \includegraphics[trim=0 105 0 0, clip, width=0.95\textwidth]{figures/root2}
-    \caption{\label{fig:root2}}
-\end{subfigure}
-\caption{Detail of 3D printed wing \subref{fig:wing1} and wing root \subref{fig:root2} design
-  showing concentric-cylindrical design to allow passive pivoting of the wing surface.}
-\end{figure}
-
-\subsection*{Measurement apparatus}
-The apparatus used to measure the forces and kinematics of flapping wings was
-based on a FUTEK LSM250 parallelogram load cell with a USB interface for data
-logging.  This force measurement system was capable of a sampling rate of 1 kHz
-and a maximum measurement precision of 1\e{-9} N.  Mounted directly to the load
-cell was a wing-flapping mechanism, which consisted of a DC gear motor mated to
-a crank and connecting rod used to drive the experimental wings in a roughly
-sinusoidal flapping motion. This system is pictured in
-Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency of
-the load cell and wing driving mechanism was roughly 70 Hz, the natural
-frequency for the first bending mode of the experimental wing was approximately
-26 Hz, and experimental wings were flapped at or below approximately 3.5 Hz.
-All experimental force measurements were filtered using a high-order zero-phase
-digital low-pass filter with a cutoff frequency of 7.5 times the flapping
-frequency, which preserved the primary inertial and aerodynamic forces acting on
-the wings while eliminating the faster vibrations of the wing structure and the
-load cell apparatus.
-
-% % See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
-\begin{figure}%[!htb]
-\centering
-\begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \includegraphics[height=5cm]{figures/apparatus2}
-    \caption{\label{fig:apparatus2}}
-\end{subfigure}
-\begin{subfigure}[t]{0.45\textwidth}
-    \centering
-    \includegraphics[height=5cm]{figures/apparatus3}
-    \caption{\label{fig:apparatus3}}
-\end{subfigure}
-\caption{\subref{fig:apparatus2} Front view of the flapping wing apparatus
-  showing the wing mounted on its driving mechanism attached to the force
-  sensor. \subref{fig:apparatus3} Close-up view of the 3D printed frame and
-  crank mechanism used to drive the flapping motion of the wing.}
-\end{figure}
-
-\subsection*{High-speed measurement of flapping kinematics}
-A Casio EXILIM EX-FH20 color video camera was used to record each experiment at
-210 frames per second from a top-down perspective as well as a side view
-provided by a mirror mounted in the field of view.  The main beam of each wing
-was colored red to enable precise tracking of the stroke angle.  The top-down
-projected area of the wing was used to extract deflection angle, and tracking of
-the wing tip in the side-view was used to extract the angle of deviation out of
-the wing's stroke plane.  These three angles precisely define the configuration
-of the wing.  A schematic diagram of this apparatus is shown in
-Figure~\ref{fig:apparatus}. A variety of image-processing techniques were used
-to extract the kinematics of the wing motion from high-speed video.  The
-red-colored main beam was identified with a color filter and a line was fitted
-using a Hough Transform to identify the stroke angle.  The wing deflection angle
-was computed by measuring the projected area of the wing in the image and
-compared with the known dimensions of the rectangular wing shape. The
-out-of-plane deviation angle of the wing was measured by tracking the leading
-and trailing edge corners of the wingtip in the side-view of the mirror. We
-refer the reader to our previous work for a more detailed description of our
-methods of extracting wing kinematics from high speed video and processing the
-measured data\cite{richter2012thesis}.
-% Figure~\ref{fig:kinematics} provides sample force data and measured wing
-% kinematics from one experiment.
-
-% \begin{figure}[ht]
-% \centering
-% \includegraphics[width=\linewidth]{figures/stream}
-% \caption{Legend (350 words max). Example legend text.}
-% \label{fig:stream}
-% \end{figure}
-
-% \begin{figure}
-% \centering
-% \begin{subfigure}[b]{.45\linewidth}
-% \includegraphics[width=\linewidth]{figures/regforce1}
-% \caption{Measured vertical force.}
-% \label{fig:regforce}
-% \end{subfigure}
-% \begin{subfigure}[b]{.45\linewidth}
-% \includegraphics[width=\linewidth]{figures/regstroke1}
-% \caption{Wing stroke angle.}
-% \label{fig:regstroke}
-% \end{subfigure}
-
-% \begin{subfigure}[b]{.45\linewidth}
-% \includegraphics[width=\linewidth]{figures/regdef1}
-% \caption{Wing deflection angle.}
-% \label{fig:regdef}
-% \end{subfigure}
-% \begin{subfigure}[b]{.45\linewidth}
-% \includegraphics[width=\linewidth]{figures/regdev1}
-% \caption{Wing deviation angle.}
-% \label{fig:regdev}
-% \end{subfigure}
-% \caption{Measured \subref{fig:regforce} vertical force, \subref{fig:regstroke} wing stroke angle,
-%   \subref{fig:regdef} deflection angle, \subref{fig:regdev} and deviation angle from one experiment,
-%   registered and averaged over 18 flapping periods.}
-% \label{fig:kinematics}
-% \end{figure}
-
 \bibliographystyle{abbrv}

### 2021-01-19 paper/main.tex
Edit Methods text

@@
 vertical force sensor, with a high-speed video camera recording the motion of
-the wing from a top-down view and (using a mirror) also from a side-view.
-Figure~\ref{fig:apparatus} provides a diagram of the apparatus as well as
-illustrations of the three primary variables describing the orientation of the
-wing in the apparatus: Stroke ($\phi$), deflection ($\psi$) and deviation
-($\theta$). The stroke motion is enabled by a hinge joint and is the only
-actuated degree of freedom in the system. The deflection of the wing is enabled
-by a hinge in the root of the wing, and is a passive motion resulting from both
-inertial and aerodynamic effects. Deviation refers to the passive vertical
-motion of the wing out of its horizontal flapping plane, and is enabled only by
-the tolerances of the hinges and the flexibility of the wing structure itself.
+the wing from a top-down view and (using a mirror) also from a side-view. The
+top-down and side views were used to detect and track the orientation of the
+wing.  Figure~\ref{fig:apparatus} provides a diagram of the apparatus as well as
+illustrations of the three primary variables describing the wing orientation:
+Stroke ($\phi$), deflection ($\psi$) and deviation ($\theta$). The stroke motion
+is enabled by a hinge joint and is the only actuated degree of freedom in the
+system. The deflection of the wing is enabled by a hinge in the root of the
+wing, and is a passive motion resulting from both inertial and aerodynamic
+effects. Deviation refers to the passive vertical motion of the wing out of its
+horizontal flapping plane, and is enabled only by the tolerances of the hinges
+and the flexibility of the wing structure itself. These three variables, and
+their derivatives, are some of the inputs required to model lift forces. The
+other inputs are the geometric and inertial properties of the wings, obtained
+from CAD designs and physical measurement. Next, we describe the wing design and
+fabrication, the force measurement apparatus, and the method of extracting wing
+kinematics from high-speed video.
 

### 2021-02-01 paper/main.tex
Add color package
Add highlighted notes
Move highlighted comments to TODOs in the .tex file
Replace Whitney-Wood and Pesavento-Want with initials and references

@@
 \usepackage{fullpage}
+\usepackage{color}
 
@@
 
-% % See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
+% TODO: Mention empirically measuring impulse response.
+
+% See https://tex.stackexchange.com/questions/286737/sub-figures-of-different-sizes-grid-layout
 \begin{figure}%[!htb]
@@
 measured data\cite{richter2012thesis}.
+
+% TODO: Mention polynomial fitting then differentiation to get derivatives.
+
 % Figure~\ref{fig:kinematics} provides sample force data and measured wing
@@
 
+% TODO: Make sure to define all variables used in these tables.
+% TODO: Figure out why we're calling it "U-squared".
 \begin{table}[ht]
@@
 \(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.28\e{-3}\\
-Whitney-Wood & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.21\e{-3}\\
-Pesavento-Wang & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.20\e{-3}\\
+WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.21\e{-3}\\
+PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.20\e{-3}\\
 \hline
@@
 
+% TODO: Make sure the bibliography entries have the right elements and are capitalized properly.
 \bibliographystyle{abbrv}

### 2021-02-01 code/modeleval_rsi21.m

@@ -214,9 +214,9 @@ text(sz_plti,mean_nofit_mae_plti,'   U^2 Planar 2','rotation',rotangle,'fontsize
 plot(sz_rw_liftonly,mean_nofit_mae_rw_liftonly,'c.','Markersize',largesize);
 text(sz_rw_liftonly,mean_nofit_mae_rw_liftonly,'   U^2 Total','rotation',rotangle,'fontsize',fsize);
 plot(sz_rw_liftam,mean_nofit_mae_rw_liftam,'r.','Markersize',largesize);
-text(sz_rw_liftam,mean_nofit_mae_rw_liftam,'   Whitney-Wood','rotation',rotangle,'fontsize',fsize);
+text(sz_rw_liftam,mean_nofit_mae_rw_liftam,'   WW','rotation',rotangle,'fontsize',fsize);
 plot(sz_jw,mean_nofit_mae_jw,'m.','Markersize',largesize);
-text(sz_jw,mean_nofit_mae_jw,'   Pesavento-Wang','rotation',rotangle,'fontsize',fsize);
+text(sz_jw,mean_nofit_mae_jw,'   PW','rotation',rotangle,'fontsize',fsize);
 plot(sz_eq1,mean_nofit_mae_eq1,'k.','Markersize',smallsize);
 plot(sz_eq2,mean_nofit_mae_eq2,'k.','Markersize',smallsize);
 plot(sz_eq3,mean_nofit_mae_eq3,'k.','Markersize',smallsize);
@@ -227,7 +227,7 @@ ylabel('Mean Absolute Error (10^-^3 N)','fontsize',fsize)
 xlabel('Equation Size (number of operators)','fontsize',fsize)
 
 xmin = 0;
-xmax = 180;
+xmax = 150;
 ymin = 0;
 ymax = 1.9e-3;
 xlim([xmin xmax])
@@ -337,9 +337,9 @@ text(sz_plti,mean_fit_mae_plti,'   U^2 Planar 2','rotation',rotangle,'fontsize',
 plot(sz_rw_liftonly,mean_fit_mae_rw_liftonly,'c.','Markersize',largesize);
 text(sz_rw_liftonly,mean_fit_mae_rw_liftonly,'   U^2 Total','rotation',rotangle,'fontsize',fsize);
 plot(sz_rw_liftam,mean_fit_mae_rw_liftam,'r.','Markersize',largesize);
[…]

### 2021-02-01 paper/main.tex
Add Fitted MAE columm

@@
 \begin{table}[ht]
+\small
 \centering
@@
 \hline
-  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & {\bf MAE (N)}\\
+  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & {\bf Fitted MAE (N)} & {\bf MAE (N)}\\
 \hline
-\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 1.19\e{-3}\\
-\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 1.18\e{-3}\\
-\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 1.14\e{-3}\\
-\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 1.14\e{-3}\\
-\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 1.12\e{-3}\\
+\(EQ _5\) & \(C_1RY_{CM}\dot{\psi}^2\ + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 12 & 8.79\e{-4} & 1.19\e{-3}\\
+\(EQ _4\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 14 & 8.66\e{-4} & 1.18\e{-3}\\
+\(EQ _3\) & \(C_1RY_{CM}\dot{\psi}^2\cos(\psi-\psi^2) + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\) & 16 & 8.49\e{-4} & 1.14\e{-3}\\
+\(EQ _2\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\dot{\phi}\dot{\psi}\psi - C_4R\ddot{\theta}\) & 19 & 8.15\e{-4} & 1.14\e{-3}\\
+\(EQ _1\) & \(C_1RY_{CM}\dot{\psi}^2\cos\psi + C_2RX_{CM}\dot{\phi}\psi - C_3R\ddot{\theta}\dot{\phi}\psi - C_4R\dot{\phi}\dot{\psi}\psi - C_5R\ddot{\theta}\) & 24 & 7.73\e{-4} & 1.12\e{-3}\\
 \hline
-\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.46\e{-3}\\
-\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.28\e{-3}\\
-\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.28\e{-3}\\
-WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.21\e{-3}\\
-PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.20\e{-3}\\
+\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
+\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
+\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
+WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
+PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
 \hline
@@
 \end{table}
+
+% TODO: Remove these:
+% Analytical models for publication
+% mean_nofit_mae_plpi =    1.4681e-03
+% mean_nofit_mae_plti =    1.2881e-03
+% mean_nofit_mae_rw_liftonly =    1.2870e-03
+% mean_nofit_mae_rw_liftam =    1.2281e-03
+% mean_nofit_mae_jw =    1.1985e-03
+% Eureqa models
+% mean_nofit_mae_eq1 =    1.1153e-03
+% mean_nofit_mae_eq2 =    1.1429e-03
+% mean_nofit_mae_eq3 =    1.1443e-03
+% mean_nofit_mae_eq4 =    1.1769e-03
+% mean_nofit_mae_eq6 =    1.1909e-03
+% Fitting and testing on testing data...
+% Analytical models for publication
+% mean_fit_mae_plpi =    1.3793e-03
+% mean_fit_mae_plti =    1.1836e-03
+% mean_fit_mae_rw_liftonly =    1.1833e-03
+% mean_fit_mae_rw_liftam =    1.0204e-03
+% mean_fit_mae_jw =    1.0088e-03
+% Eureqa models
+% mean_fit_mae_eq1 =    7.7317e-04
+% mean_fit_mae_eq2 =    8.1508e-04
+% mean_fit_mae_eq3 =    8.4913e-04
+% mean_fit_mae_eq4 =    8.6626e-04
+% mean_fit_mae_eq6 =    8.7928e-04
 

### 2021-02-01 paper/main.tex
Label columns as fitted and not fitted
Add TODO
Explain the difference between fitted and not fitted results

@@
 
+% TODO: Reword this section to not focus on the number 58 right off the bat.
 \subsection*{Data-driven models}
@@
 describe the wing's shape and distribution of area, totaling 31 separate
-variables.
+variables. By training on 58 experiments simultaneously, the symbolic regression
+algorithm generated equations that fit all of the data, using the some, but not
+all, of the provided variables to account for variations in kinematics and wing
+geometry between experiments. The search was conducted for approximately 24
+hours, using 60 cores on a computing cluster.
 
-By training on 58 experiments simultaneously, the symbolic regression algorithm
-generated equations that fit all of the data, using the some, but not all, of
-the provided variables to account for variations in kinematics and wing geometry
-between experiments.   Table~\ref{table:eqmodels} lists five of the leading
-candidate models, which were trained using mean absolute error (MAE) as the
-fitness metric during the search. The search was conducted for approximately 24
-hours, using 60 cores on a computing cluster.
+Table~\ref{table:eqmodels} lists the performance of five of the leading
+candidate models, along with five analytical models for comparison. We measure
+performance as the mean absolute error (MAE) between predicted and measured
+vertical force in the 10 test set experiments withheld from training. Our
+primary results, provided in the column labeled ``MAE Not Fitted'', use
+coefficient values for all models that were optimized on the training set.
+However, since the accuracy of the analytical models may depend on coefficients
+that vary with wing shape, flapping speed, or other factors, we also provide
+secondary results in the column labeled ``MAE Fitted'', which use coefficient
+values that were fitted to the test set experiments.
 
@@
 % TODO: Figure out why we're calling it "U-squared".
+% TODO: Decide on better terms than "fitted" and "not fitted".
 \begin{table}[ht]
@@
 \hline
-  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & {\bf Fitted MAE (N)} & {\bf MAE (N)}\\
+  {\bf Model} & {\bf Terms/Formula} & {\bf Size} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\
+    Fitted\end{tabular} & \begin{tabular}{@{}c@{}}{\bf MAE (N)} \\ Not Fitted\end{tabular}\\
 \hline
@@
 
+% TODO: Remove all occurrences of the word "pareto" from the text, figures and filenames.
 \begin{figure}

### 2021-02-07 paper/main.tex
Revisions to abstract, introduction and methods

@@
 We explore a data-driven approach to symbolically modeling the lift produced by
-the flapping wing of a micro air vehicle.  The aerodynamics of flapping flight
-pose a modeling challenge due to complex effects related to vortex dynamics,
-wake interaction, aeroelasticity and three-dimensional flow, yet simple and
-accurate equations are useful for engineering purposes and optimization of wing
-design.  Experiments were conducted to measure the flapping kinematics and
-instantaneous lift forces produced by eleven different 3D printed wing
-geometries at three different flapping amplitudes and multiple speeds, totaling
-68 distinct sets of data.  58 of these data sets were used to train data-driven
-models using a symbolic regression algorithm, while the remaining sets were
-reserved for model testing.  Data-driven models were found to be more accurate
-and much simpler than existing quasi-steady analytical models consistent with
-aerodynamic theory.  Furthermore, data-driven models were found to separate
-distinct physical effects into inertial and aerodynamic components bearing
-remarkable resemblance to their analytical counterparts. This result suggests
-that a data-driven approach could be a powerful tool in separating and
+flapping wings at the scale of a micro air vehicle.  The aerodynamics of
+flapping flight pose a modeling challenge due to complex effects related to
+vortex dynamics, wake interaction, aeroelasticity and three-dimensional flow,
+yet simple and accurate equations are useful for engineering purposes and
+optimization of wing design.  Experiments were conducted to measure the flapping
+kinematics and instantaneous lift forces produced by eleven different 3D printed
+wing geometries at three different flapping amplitudes and multiple speeds,
+totaling 68 distinct sets of data.  58 of these data sets were used to train
+data-driven models using a symbolic regression algorithm, while the remaining
+sets were reserved for model testing.  Data-driven models were found to be more
+accurate and much simpler than existing quasi-steady analytical models
+consistent with aerodynamic theory.  Furthermore, data-driven models were found
+to separate distinct physical effects into inertial and aerodynamic components
+bearing remarkable resemblance to their analytical counterparts.  This result
+suggests that a data-driven approach could be a powerful tool in separating and
 identifying the structural building blocks of a wide variety of unknown
@@
 of a wing's kinematic variables such as velocity and angle of attack.  Both
-methods have shown qualitative and quantitative agreement with experimental
-measurements \cite{wang2004unsteady}.
+methods have shown qualitative and quantitative agreement with experiments
+\cite{wang2004unsteady}.
 
@@
 of symbolic regression, which harnesses computational power to conduct an
-automated search over the space of all kinematic variables and geometric
-parameters to find equations that most accurately predict the forces on a wing.
-Symbolic regression algorithms are based on evolutionary computation
+automated search over the space of relationships between kinematic variables and
+geometric parameters to find equations that most accurately predict the forces
+on a wing.  Symbolic regression algorithms are based on evolutionary computation
 \cite{koza1992genetic,forrest1993genetic} and have been used to model complex
@@
 Our experimental apparatus consisted of a wing and actuator mounted to a
-vertical force sensor, with a high-speed video camera recording the motion of
-the wing from a top-down view and (using a mirror) also from a side-view. The
-top-down and side views were used to detect and track the orientation of the
-wing.  Figure~\ref{fig:apparatus} provides a diagram of the apparatus as well as
-illustrations of the three primary variables describing the wing orientation:
-Stroke ($\phi$), deflection ($\psi$) and deviation ($\theta$). The stroke motion
-is enabled by a hinge joint and is the only actuated degree of freedom in the
-system. The deflection of the wing is enabled by a hinge in the root of the
-wing, and is a passive motion resulting from both inertial and aerodynamic
-effects. Deviation refers to the passive vertical motion of the wing out of its
-horizontal flapping plane, and is enabled only by the tolerances of the hinges
-and the flexibility of the wing structure itself. These three variables, and
-their derivatives, are some of the inputs required to model lift forces. The
-other inputs are the geometric and inertial properties of the wings, obtained
-from CAD designs and physical measurement. Next, we describe the wing design and
-fabrication, the force measurement apparatus, and the method of extracting wing
-kinematics from high-speed video.
+vertical force sensor, with a high-speed color video camera recording the motion
+of the wing from a top-down view, and also from a side view using a mirror. The
+top-down and side views were used to detect and track the position and
+orientation of the wing.  Figure~\ref{fig:apparatus} provides a diagram of the
+apparatus as well as illustrations of the three primary variables describing the
+wing orientation: Stroke ($\phi$), deflection ($\psi$) and deviation ($\theta$).
+The stroke motion is enabled by a hinge joint and is the only actuated degree of
+freedom in the system. The deflection of the wing is enabled by a hinge in the
+root of the wing, and is a passive motion resulting from both inertial and
+aerodynamic effects. Deviation refers to the passive vertical motion of the wing
+out of its horizontal flapping plane, and is enabled only by the tolerances of
+the hinges and the flexibility of the wing structure itself. These three
+variables, and their derivatives, are some of the inputs required to model lift
+forces. The other inputs are the geometric and inertial properties of the wings,
+obtained from CAD designs and physical measurement. Next, we describe the wing
+design and fabrication, the force measurement apparatus, and the method of
+extracting wing kinematics from high-speed video.
 
@@
 \caption{Schematic diagram of the experimental apparatus
-  showing three kinematic variables of the wing motion: stroke ($\phi$),
+  showing three kinematic variables of the wing motion: Stroke ($\phi$),
   deflection ($\psi$) and deviation ($\theta$).\label{fig:apparatus}}
@@
 The experimental wings were fabricated using an Objet Connex500 3D printer with
-the rigid Objet FullCure720 material.  Flapping flight with 3D printed wings was
-reported in \cite{richter2011untethered}. The wing design consisted of an
-interlocking root and tip, allowing the wing surface to pivot naturally during
-flapping from its neutral vertical position to an angle of attack of
+the rigid Objet FullCure720 material. In prior work, we demonstrated flapping
+flight of a micro air vehicle with wings made using this material
+\cite{richter2011untethered}. The wing design consisted of an interlocking
+hinged root and tip, allowing the wing surface to pivot naturally during
+flapping from its neutral vertical position to a deflection angle of
 approximately 45\textdegree.  The wing design is illustrated in
@@
 \subsection*{Measurement apparatus}
-The apparatus used to measure the forces and kinematics of flapping wings was
-based on a FUTEK LSM250 parallelogram load cell with a USB interface for data
-logging.  This force measurement system was capable of a sampling rate of 1 kHz
-and a maximum measurement precision of 1\e{-9} N.  Mounted directly to the load
-cell was a wing-flapping mechanism, which consisted of a DC gear motor mated to
-a crank and connecting rod used to drive the experimental wings in a roughly
-sinusoidal flapping motion. This system is pictured in
-Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural frequency of
-the load cell and wing driving mechanism was roughly 70 Hz, the natural
-frequency for the first bending mode of the experimental wing was approximately
-26 Hz, and experimental wings were flapped at or below approximately 3.5 Hz.
-All experimental force measurements were filtered using a high-order zero-phase
-digital low-pass filter with a cutoff frequency of 7.5 times the flapping
-frequency, which preserved the primary inertial and aerodynamic forces acting on
-the wings while eliminating the faster vibrations of the wing structure and the
-load cell apparatus.
+The force measurement apparatus was based on a FUTEK LSM250 parallelogram load
+cell, capable of a sampling rate of 1 kHz and a precision of 1\e{-9} N.  Mounted
+directly to the load cell was a wing-flapping mechanism, which consisted of a DC
+gear motor powering a crank and connecting rod used to drive the experimental
+wings in a roughly sinusoidal flapping motion. This system is pictured in
+Figures~\ref{fig:apparatus2} and \ref{fig:apparatus3}.  The natural vibration
+frequency of the load cell and wing driving mechanism was roughly 70 Hz, the
+natural vibration frequency for the first bending mode of the experimental wing
+was approximately 26 Hz, and experimental wings were flapped at or below
+approximately 3.5 Hz.  All experimental force measurements were filtered using a
+high-order zero-phase digital low-pass filter with a cutoff frequency of 7.5
+times the flapping frequency, which preserved the primary inertial and
+aerodynamic forces acting on the wings while eliminating the faster vibrations
+of the wing structure and the load cell apparatus.
 
@@
 210 frames per second from a top-down perspective as well as a side view
-provided by a mirror mounted in the field of view.  The main beam of each wing
-was colored red to enable precise tracking of the stroke angle.  The top-down
-projected area of the wing was used to extract deflection angle, and tracking of
-the wing tip in the side-view was used to extract the angle of deviation out of
-the wing's stroke plane.  These three angles precisely define the configuration
-of the wing.  A schematic diagram of this apparatus is shown in
-Figure~\ref{fig:apparatus}. A variety of image-processing techniques were used
-to extract the kinematics of the wing motion from high-speed video.  The
-red-colored main beam was identified with a color filter and a line was fitted
-using a Hough Transform to identify the stroke angle.  The wing deflection angle
-was computed by measuring the projected area of the wing in the image and
-compared with the known dimensions of the rectangular wing shape. The
-out-of-plane deviation angle of the wing was measured by tracking the leading
-and trailing edge corners of the wingtip in the side-view of the mirror. We
-refer the reader to our previous work for a more detailed description of our
-methods of extracting wing kinematics from high speed video and processing the
-measured data\cite{richter2012thesis}.
+provided by a mirror mounted in the field of view. Both the top-down and side
+views featured matte black backdrops to facilitate easy identification of the
+wing in video frames. A variety of image-processing techniques were used to
+extract the kinematics of the wing motion from each video frame. The main beam
+of each wing was colored red and identified with a color filter in the top-down
+view. A line was then fitted using the Hough Transform to identify the stroke
+angle. The wing deflection angle was computed by measuring the projected area of
+the wing in the top-down view, calibrated using the known dimensions of the
+rectangular wing shape. The out-of-plane deviation angle of the wing was
+measured by tracking the leading and trailing edge corners of the wingtip in the
+side view and comparing their positions to the nominal position of the wingtip
+with zero deflection and deviation throughout the wing stroke. We refer the
+reader to our previous work for further details and illustrations of our methods
+for extracting wing kinematics from high speed video and processing the measured
+data~\cite{richter2012thesis}.
 

### 2021-02-07 paper/main.tex
Add subsection on data smoothing and registration

@@
 obtained from CAD designs and physical measurement. Next, we describe the wing
-design and fabrication, the force measurement apparatus, and the method of
-extracting wing kinematics from high-speed video.
+design and fabrication, the force measurement apparatus, and the methods of
+extracting and processing wing kinematics data from high-speed video.
 
@@
 data~\cite{richter2012thesis}.
+
+\subsection*{Data Smoothing and Registration}
+Stroke, deflection and deviation measurements were smoothed with a
+Savitzky–Golay filter, providing not only a reduction in noise in the signals
+themselves, but also smooth first and second derivatives of these signals. These
+kinematic variables were temporally aligned and resampled to the 1 kHz frequency
+of the force measurement. Finally, for each wing at each flapping frequency and
+amplitude, the data for 10-20 wing strokes were registered and averaged together
+to create a composite sample representing the force and kinematics of a single
+complete wing stroke of the given wing, frequency and amplitude. These composite
+samples were then used for data-driven modeling and analysis.
 

### 2021-02-07 paper/main.tex
Minor edit to first paragraph in Results section
Add TODO
Edit data-driven models subsection

@@
 \subsection*{Data-driven models}
-The strategy for modeling lift forces symbolically was to train candidate
-equations on 58 of the 68 total experiments simultaneously, with the remaining
-ten reserved for testing.  These 58 training sets were intended to represent a
-diverse sampling of wing geometry, flapping speed and amplitude combinations.
-The results from these 58 experiments were compiled into a single dataset
-listing roughly two periods of smoothed, registered flapping data per
-experiment.  The data consisted of instantaneous measured vertical force along
-with the measured value of each kinematic variable, the first and second
-derivatives of those variables, and the geometric parameters necessary to
-describe the wing's shape and distribution of area, totaling 31 separate
-variables. By training on 58 experiments simultaneously, the symbolic regression
-algorithm generated equations that fit all of the data, using the some, but not
-all, of the provided variables to account for variations in kinematics and wing
-geometry between experiments. The search was conducted for approximately 24
-hours, using 60 cores on a computing cluster.
+We trained a set of data-driven models to predict the instantaneous vertical
+force produced by flapping wings as a function of the kinematic variables and
+derivatives describing their motion, and geometric parameters describing the
+wing shape, totaling 31 separate variables. We trained these models on data from
+58 experiments, selected at random from our 68 total experiments, which
+represent a diverse sampling of wing geometry, flapping speed and amplitude
+combinations. The remaining 10 experiments were reserved for testing. Models
+were trained using the symbolic regression algorithm developed by Schmidt and
+Lipson \cite{schmidt2009distilling} to minimized the mean absolute error (MAE)
+between measured and predicted vertical force. The training process ran for
+approximately 24 hours using 60 cores on a computing cluster.
 
+By training on data representing a range of wing shapes and flapping speeds and
+amplitudes, the model search effectively narrowed the large set of input
+variables to the most preditive information and only utilized those terms that
+accounted for variations in kinematics and wing geometry between experiments.
 Table~\ref{table:eqmodels} lists the performance of five of the leading
-candidate models, along with five analytical models for comparison. We measure
-performance as the mean absolute error (MAE) between predicted and measured
-vertical force in the 10 test set experiments withheld from training. Our
-primary results, provided in the column labeled ``MAE Not Fitted'', use
-coefficient values for all models that were optimized on the training set.
-However, since the accuracy of the analytical models may depend on coefficients
-that vary with wing shape, flapping speed, or other factors, we also provide
-secondary results in the column labeled ``MAE Fitted'', which use coefficient
-values that were fitted to the test set experiments.
+candidate models, along with five analytical models for comparison, as measured
+by MAE in the 10 test set experiments withheld from training. These five models
+were selected from the Pareto frontier between MAE and formula complexity when
+training was terminated. The data-driven models presented here are functions of
+the kinematic variables ($\phi$, $\psi$, and $\theta$) and their derivatives,
+the wing span ($R$), the span-wise and chord-wise locations of the center of
+wing area ($X_{CM}$ and $Y_{CM}$), and constant coefficients. The analytical
+models for comparison will be described in the following section.
+
+Our primary results, provided in the column labeled ``MAE Not Fitted'', use
+fixed coefficient values for all models that were jointly optimized on the
+training set.  However, since the accuracy of the analytical models may depend
+on coefficients that vary with wing shape, flapping speed, or other factors, we
+also provide secondary results in the column labeled ``MAE Fitted'', which use
+coefficient values that were fitted individually to each of the test set
+experiments. In both categories of results, the data-driven models are more
+accurate, demonstrating that the relative performance of data-driven models is
+not simply due to the choice or optimization of coefficient values.
 
@@
 % TODO: Figure out why we're calling it "U-squared".
+% TODO: If U-squared is "velocity squared" then  make sure that's consistent
+% with the use of $v^2$ in the L_{trans} formula (equation 1) and elsewhere.
 % TODO: Decide on better terms than "fitted" and "not fitted".
+% TODO: Explain that the 5 EQ models were taken from the Pareto frontier of
+% complexity vs. accuracy at the time of convergence.
+% TODO: Explain that C_1,...,C_5 are not the same in each EQ. Maybe do C_i,_j
 \begin{table}[ht]

### 2021-02-07 paper/main.tex
Add TODOs
Use V for velocity (including U^2 -> V^2)

@@
 % TODO: Make sure to define all variables used in these tables.
-% TODO: Figure out why we're calling it "U-squared".
+% TODO: Figure out why we're calling it "U-squared". -->> Change to V^2
+% everywhere (see Whitney Wood paper just below equation 2.18).
 % TODO: If U-squared is "velocity squared" then  make sure that's consistent
@@
 \hline
-\(U^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
-\(U^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
-\(U^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
+\(V^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
+\(V^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
+\(V^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
 WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
@@
 \begin{equation}
-L_{trans}=(\nicefrac{1}{2})\rho v^2AC_L,
+L_{trans}=(\nicefrac{1}{2})\rho V^2AC_L,
 \end{equation}
-where \(v\) is the wing velocity, \(\rho\) is the fluid density, \(A\) is the area of the wing surface, and \(C_L\) is the coefficient of lift.
+where \(V\) is the wing velocity, \(\rho\) is the fluid density, \(A\) is the area of the wing surface, and \(C_L\) is the coefficient of lift.
 The theoretical value of the coefficient of lift depends on the angle of attack \(\alpha\):
@@
 \end{equation}
+
+% TODO: Understand and verify (numerically) how we get an effective v^2 via
+% integration along the span, and document that in the code.
 

### 2021-02-07 code/modeleval_rsi21.m

@@ -208,11 +208,11 @@ set(f1,'units','centimeters','paperunits','centimeters','papersize',[width heigh
     'paperposition',[0 0 width height])
 
 plot(sz_plpi,mean_nofit_mae_plpi,'b.','Markersize',largesize);
-text(sz_plpi,mean_nofit_mae_plpi,'   U^2 Planar 1','rotation',rotangle,'fontsize',fsize);
+text(sz_plpi,mean_nofit_mae_plpi,'   V^2 Planar 1','rotation',rotangle,'fontsize',fsize);
 plot(sz_plti,mean_nofit_mae_plti,'g.','Markersize',largesize);
-text(sz_plti,mean_nofit_mae_plti,'   U^2 Planar 2','rotation',rotangle,'fontsize',fsize);
+text(sz_plti,mean_nofit_mae_plti,'   V^2 Planar 2','rotation',rotangle,'fontsize',fsize);
 plot(sz_rw_liftonly,mean_nofit_mae_rw_liftonly,'c.','Markersize',largesize);
-text(sz_rw_liftonly,mean_nofit_mae_rw_liftonly,'   U^2 Total','rotation',rotangle,'fontsize',fsize);
+text(sz_rw_liftonly,mean_nofit_mae_rw_liftonly,'   V^2 Total','rotation',rotangle,'fontsize',fsize);
 plot(sz_rw_liftam,mean_nofit_mae_rw_liftam,'r.','Markersize',largesize);
 text(sz_rw_liftam,mean_nofit_mae_rw_liftam,'   WW','rotation',rotangle,'fontsize',fsize);
 plot(sz_jw,mean_nofit_mae_jw,'m.','Markersize',largesize);
@@ -331,11 +331,11 @@ set(f2,'units','centimeters','paperunits','centimeters','papersize',[width heigh
     'paperposition',[0 0 width height])
 
 plot(sz_plpi,mean_fit_mae_plpi,'b.','Markersize',largesize);
-text(sz_plpi,mean_fit_mae_plpi,'   U^2 Planar 1','rotation',rotangle,'fontsize',fsize);
+text(sz_plpi,mean_fit_mae_plpi,'   V^2 Planar 1','rotation',rotangle,'fontsize',fsize);
 plot(sz_plti,mean_fit_mae_plti,'g.','Markersize',largesize);
[…]

### 2021-02-07 paper/main.tex
Clarify analytical model names and terms

@@
 \hline
-\(V^2\) Planar 1 & \(L_{planar},  F_{inertial\:(pendulum\:only)}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
-\(V^2\) Planar 2 & \(L_{planar}, F_{inertial}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
-\(V^2\) Total & \(L_{trans}, F_{inertial}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
-WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
-PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans}, AM_{trans-rot}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
+\(V^2\) Planar 1 & \(L_{planar},  F_{pendulum}\) & 32 & 1.38\e{-3} & 1.47\e{-3}\\
+\(V^2\) Planar 2 & \(L_{planar}, F_{pendulum}, F_{deviation}\) & 46 & 1.18\e{-3} & 1.29\e{-3}\\
+\(V^2\) Total & \(L_{trans}, F_{pendulum}, F_{deviation}\) & 96 & 1.18\e{-3} & 1.29\e{-3}\\
+WW~\cite{whitney2010aeromechanics} & \(L_{trans}, AM_{trans}, AM_{rot}, F_{pendulum}, F_{deviation}\) & 121 & 1.02\e{-3} & 1.23\e{-3}\\
+PW~\cite{pesavento2004falling} & \(L_{trans}, L_{rot}, AM_{trans},
+    AM_{trans-rot}, F_{pendulum}, F_{deviation}\) & 130 & 1.01\e{-3} & 1.20\e{-3}\\
 \hline
@@
 \end{equation}
+In Table~\ref{table:eqmodels} and Figure~\ref{fig:pareto_mae}, we present
+results for three models whose only aerodynamic component is translational lift
+proportional to wing velocity, which we term $V^2$ Planar 1, $V^2$ Planar 2, and
+$V^2$ Total.
 
@@
 
-Other aerodynamic contributors include rotational lift proportional to the
+Additional aerodynamic effects include rotational lift proportional to the
 product of a wing's rotational and translational velocities, labeled $L_{rot}$,
-as well as "added-mass" effects related to the accelerations of fluid
+as well as ``added-mass'' effects related to the accelerations of fluid
 surrounding the wing \cite{sedov1965two}.  These added mass terms are
 proportional to the wing's linear acceleration ($AM_{trans}$), rotational
-acceleration ($AM_{rot}$), or a product of the two ($AM_{trans-rot}$).  Two sets
-of quasi-steady equations from the literature include various combinations of
-such terms \cite{whitney2010aeromechanics,pesavento2004falling}, which allow
-more flexible model fitting through additional coefficients.
+acceleration ($AM_{rot}$), or a product of the two ($AM_{trans-rot}$). Whitney
+and Wood~\cite{whitney2010aeromechanics} and Pesavento and
+Wang~\cite{pesavento2004falling} describe quasi-steady models that use different
+combinations of such terms. We refer to these models in
+Table~\ref{table:eqmodels} and Figure~\ref{fig:pareto_mae} as WW and PW, respectively.
 
@@
 secondary contribution results from small angle deviations from the horizontal
-flapping plane ($\theta$).  These angles are depicted in
-Figure~\ref{fig:apparatus}.  These inertial forces are given by:
+flapping plane ($\theta$). These inertial forces are given by:
 \begin{equation}\label{sdbla}
@@
 \end{equation}
-\noindent where \(X_{CM}\) and \(Y_{CM}\) are the coordinates of the center of wing area and \(M\) is the mass of the wing surface.
+\noindent where \(X_{CM}\) and \(Y_{CM}\) are the span-wise and chord-wise
+locations of the center of wing area and \(M\) is the mass of the wing surface.
+All of our analytical models for comparison include $F_{pendulum}$ and all
+except $V^2$ Planar 1 include $F_{deviation}$.
 

### 2021-02-07 paper/main.tex
Edit evaluation of model performance section

@@
 study, and indicates that the analytical models are all significantly more
-complex than our data-driven models.  One of the reasons for high complexity in
-the analytical models is the dependence on wing geometry and integration of
-forces over the span and chord of the wing.  A second reason for the complexity
-in analytical models is the number of geometric and coordinate transformations
-required to compute the necessary linear and rotational velocities.  In this
-sense, several terms generated during symbolic regression can be considered to
-be approximations of much more subtle calculations of kinematic quantities.
-
-The accuracy of each equation was measured as the mean absolute error (MAE)
-between the predicted and actual vertical force (in Newtons) produced by the
-flapping wing.  Figure~\ref{fig:pareto_mae} shows the Pareto plot of all models
-considered, depicting accuracy as a function of model complexity. It is
-important to note that all data-driven models have between three and five
-coefficients, so tuning these coefficients allows further minimization of error
-compared to the analytical models, which have between one and three.
-Nevertheless, discovery of models with smaller error than theoretical
-predictions that are simultaneously much simpler is a significant feat for the
-data mining process and symbolic regression.
+complex than our data-driven models.  Figure~\ref{fig:pareto_mae} illustrates
+model accuracy as a function of complexity.  One of the reasons for high
+complexity in the analytical models is the dependence on wing geometry and
+integration of forces over the span and chord of the wing.  A second reason for
+the complexity in analytical models is the number of geometric and coordinate
+transformations required to compute the necessary linear and rotational
+velocities.  In this sense, several terms generated during symbolic regression
+can be considered to be approximations of much more subtle calculations of
+kinematic quantities. Nevertheless, discovery of models with smaller error than
+theoretical predictions that are simultaneously much simpler is a significant
+feat for the data mining process and symbolic regression.
 

### 2021-02-07 paper/main.tex
Minor changes

@@
   compute a force prediction from kinematic and geometric data.  MAE reports
-  average mean absolute error (in Newtons of vertical force) for the ten test
+  average mean absolute error (in Newtons of vertical force) for the 10 test
   experiments.}
@@
 occurs when the deflection angle ($\psi$) flips at stroke reversal.  The
-secondary contribution results from small angle deviations from the horizontal
-flapping plane ($\theta$). These inertial forces are given by:
+secondary contribution results from small angle deviations ($\theta$) from the horizontal
+flapping plane. These inertial forces are given by:
 \begin{equation}\label{sdbla}
@@
 locations of the center of wing area and \(M\) is the mass of the wing surface.
-All of our analytical models for comparison include $F_{pendulum}$ and all
+All of our analytical models for comparison include $F_{pendulum}$, and all
 except $V^2$ Planar 1 include $F_{deviation}$.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{figures/mae_nofit}
+\includegraphics[width=0.6\textwidth]{figures/mae_nofit}
 \caption{Pareto plot of all data-driven (``EQ Models'') and analytical models

### 2021-02-07 paper/main.tex
Edit generalization of data-driven models section

@@
 
-\subsection*{Performance of models trained on one experiment versus many experiments}
-Figure~\ref{fig:eq_goodbad_1} illustrates the predictive performance of two
-models on data within their training sets, while
-Figures~\ref{fig:eq_goodbad_2},~\ref{fig:eq_goodbad_3},
-and~\ref{fig:eq_goodbad_4} illustrate the performance of those same two models
-on data outside their training sets. The differences between these two models
-result from the fact that one of them, $EQ_4$ from Table~\ref{table:eqmodels}
-(drawn in blue), was trained on 58 separate experiments simultaneously, whereas
-the other one, shown below (drawn in red), is one of many models trained only on
-the single experiment shown in Figure~\ref{fig:eq_goodbad_1}:
+\subsection*{Generalization of Data-Driven Models}
+To evaluate the generalization properties of our data-driven models, we compare
+one of our models, $EQ_4$, which was trained on 58 training experiments, to a
+model that was trained on only a single experiment:
 \begin{align}
@@
 \end{align}
-Thus, $EQ_4$ shows strong qualitative and quantitative agreement with all of the
+Figure~\ref{fig:eq_goodbad_1} illustrates the predictive performance of these
+two models on data within their training sets, and indeed this figure
+illustrates the \emph{only} data used to train equation~\eqref{eq:badeq}.
+Figures~\ref{fig:eq_goodbad_2},~\ref{fig:eq_goodbad_3},
+and~\ref{fig:eq_goodbad_4} illustrate the performance of these same two models
+on test data outside their training sets.
+
+$EQ_4$ shows strong qualitative and quantitative agreement with all of the
 experiments in Figure~\ref{fig:eq_goodbad}, indicating that this model is not
-over-fit but in fact generalizes well to new data. Equation~\eqref{eq:badeq}, on
-the other hand, fails to accurately predict the recorded data outside its
+over-fit but in fact generalizes well to new data. On the other hand,
+equation~\eqref{eq:badeq} fails to accurately predict the data outside its
 training set.  Insufficient data and over-fitting are potential pitfalls of
 data-driven modeling, but are overcome in this case with a diverse sampling of
-training data.
+training data that span the parameter space of interest for the modeling task.
 
@@
 physical processes.  $EQ_4$, however, must capture some of the underlying
-structure of the physics in order to generalize to data upon which it was not
-trained.  In data-driven modeling, training on many experiments simultaneously
+structure of the physics in order to generalize to data outside its training
+set. In data-driven modeling, training on many experiments simultaneously
 helps to reject candidate equations that may fit some, but not all, of the data,
 and helps to retain those models that reflect the underlying physics and offer
-insight into the structure of the additive components that compose the observed
+insight into the structure of the components that make up the observed
 behavior of a complex dynamical system.

### 2021-02-07 paper/main.tex
Edit caption

@@
 \caption{Comparison of force predictions between $EQ_4$ (blue) and
-Equation~\eqref{eq:badeq} (red). $EQ_4$ was trained on 58 distinct experiments
-simultaneously, including the data shown in~\subref{fig:eq_goodbad_1}, whereas
-Equation~\eqref{eq:badeq} was trained \emph{only} on the data shown
-in~\subref{fig:eq_goodbad_1}.  \subref{fig:eq_goodbad_1} shows that both models
-perform well on this data within their training sets.  However, when presented
-with test data \subref{fig:eq_goodbad_2}, \subref{fig:eq_goodbad_3} and
-\subref{fig:eq_goodbad_4} not contained within either model's training set, only $EQ_4$ is able to
-generalize and provide accurate predictions, while Equation~\eqref{eq:badeq}
-fails to do so.}
+  Equation~\eqref{eq:badeq} (red). The measured force is illustrated in black.
+  $EQ_4$ was trained on 58 distinct experiments simultaneously, including the
+  data shown in~\subref{fig:eq_goodbad_1}, whereas equation~\eqref{eq:badeq} was
+  trained \emph{only} on the data shown in~\subref{fig:eq_goodbad_1}. Test
+  experiments \subref{fig:eq_goodbad_2}, \subref{fig:eq_goodbad_3} and
+  \subref{fig:eq_goodbad_4}, not contained in either model's training set, show
+  that $EQ_4$ is able to generalize and provide accurate predictions, while
+  Equation~\eqref{eq:badeq} fails to do so.}
 \label{fig:eq_goodbad}

### 2021-02-07 paper/main.tex
Add TODO
Capitalize v in v-squared expressions

@@
 
+% TODO: Make correspondence more explicit. For example, mass M is a direct
+% stand-in for wingspan R with a coefficient.
+
 \begin{table}[ht]
@@
 \hline
-Aerodynamic Lift & \(\nicefrac{1}{2}\rho v^2AC_{Lmax}\sin(2\alpha)\) & \(C_2RX_{CM}\psi\dot{\phi}\)\\
+Aerodynamic Lift & \((\nicefrac{1}{2})\rho V^2AC_{Lmax}\sin(2\alpha)\) & \(C_2RX_{CM}\psi\dot{\phi}\)\\
 Pendulum 1 & \(M\cdot Y_{CM}\ddot{\psi}\sin{\psi}\) & \it{None} \\
@@
 for why \(\psi\dot{\phi}\) is used as an approximation for an aerodynamic lift
-term proportional to \(v^2\).  These correlations between variables are
+term proportional to \(V^2\).  These correlations between variables are
 inherent in highly symmetric, periodic phenomena such as flapping flight.  In

### 2021-02-07 paper/main.tex
Minor edit

@@
 parameters in the 58 sets of training data, any candidate models that
-incorrectly mix or couple two separate physical effects are discarded upon
-testing on a data set with different parameters.
+incorrectly mix or couple two separate physical effects are discarded when
+evaluated on data representing different parameters.
 

### 2021-02-07 paper/main.tex
Minor edit

@@
 inherent in highly symmetric, periodic phenomena such as flapping flight.  In
-principle, additional experiments could be performed on flapping surfaces that
+principle, additional experiments could be performed on aerodynamic systems that
 are not bound to the natural kinematics of flapping wings to explore the full

### 2021-04-05 paper/main.tex
Add wing frames image and edit introduction to include it

@@
 are of crucial importance for flight, yet they are subtle and difficult to
-model.  Many studies have developed equations and solution techniques to
-understand the fluid flow around these wings, to compute control actions
+model. In many cases, the flapping motion is only partially actuated and
+controlled, while essential characteristics like angle of attack through the
+wing stroke result passively from complex inertial and aerodynamic forces and
+rely on the compliance of a flexible wing structure.
+Figure~\ref{fig:wing_frames} shows a sequence of frames from a high-speed video
+of a flexible 3D printed wing being actuated only in the primary flapping
+motion, yet the wing passively twists and deflects to achieve an angle of attack
+of approximately 45\textdegree, which allows it to produce vertical lift.
+
+Many studies have developed equations and solution techniques to
+understand the fluid flow around flapping wings, to compute control actions
 required for stable flight, or to optimize wing design. Aerodynamic studies
@@
 % \subsection{Automated modeling}
+
+\begin{figure}
+\centering
+\includegraphics[trim=0 120 70 170, clip, width=0.8\textwidth]{figures/wing_frames}
+\caption{Composite image showing the flapping motion of a 3D printed
+  wing captured in high-speed video.\label{fig:wing_frames}}
+\end{figure}
 
